### 2014-04-17 paper/kepler-selection.tex
Initial import of paper.
Ilya has convinced me to try submitting to Nature....  We'll see where it goes.

\documentclass{nature}

\usepackage{hyperref}
\usepackage{amsmath}
\usepackage{amssymb}
\usepackage{xcolor}

\bibliographystyle{naturemag}

\newcommand{\etaearth}{\eta_\oplus}
\newcommand{\Rpeak}{R_\mathrm{peak}}
\newcommand{\REarth}{R_\oplus}

\newcommand{\email}[1]{\href{mailto:#1}{\nolinkurl{#1}}}

\newcommand{\Will}[1]{\textcolor{cyan}{Will: #1}}

\newcommand{\apj}{The Astrophysical Journal}
\newcommand{\apjs}{The Astrophysical Journal Supplement}

\begin{document}
\title{The Occurrence of Earth-Like Planets Around Other Stars}

\author{Will M. Farr$^{1,2}$, Chris Aldridge$^{1,3}$, Kirsty Stroud$^{1,4}$ \& Ilya Mandel$^{1,5}$}

\maketitle

\begin{affiliations}
\item School of Physics and Astronomy, Univeristy of Birmingham, Birmingham, B15 2TT, United Kingdom
\item \email{w.farr@bham.ac.uk}
\item \email{CXA064@bham.ac.uk}
\item \email{KLS081@bham.ac.uk}
\item \email{ilyamandel@chgk.info}
\end{affiliations}

\begin{abstract}
  The quantity $\etaearth$, the number density of planets per star per
  logarithmic planetary radius per logarithmic orbital period,
  describes the occurrence of Earth-like extrasolar planets.
  Measurement of $\etaearth$ is complicated by the difficulty of
  detecting Earth-like planets in Earth-like orbits about Sun-like
  stars; previous estimates place $1\% \lesssim \etaearth \lesssim
  34\%$, relying on extrapolations from regions in period-radius space
  where detection is easier and the sample of observed planets is
  complete\cite{2011ApJ...738...81W,2011ApJ...738..151C,2012ApJ...745...20T},
  or modelling the selection effects carefully but using simpler
  models of the period-radius distribution of
  exoplanets\cite{2013ApJ...778...53D}, possibly using a low-noise
  subset of the available data\cite{2013PNAS..11019273P}.
  Additionally, no previous estimates account for contamination from
  false-positive planetary identifications.  Here we present
  constraints on $\etaearth$ from a parameterised forward model of the
  (correlated) period-radius distribution and the observational
  selection function in the most recent (Q17) data release from the
  Kepler
  satellite\cite{2010Sci...327..977B,2011ApJ...736...19B,2013ApJS..204...24B},
  including an empirical parameterisation for the distribution of
  false-positives in the data set.  We find $\etaearth =
  4.1^{+2.3}_{-1.7}\%$ (90\% CL).  Additionally, we find that each
  star hosts $4.04_{-0.68}^{+0.85}$ planets with $P \lesssim 3
  \mathrm{yr}$ and $R \gtrsim 0.2 \REarth$, that the peak of the
  planet-radius distribution lies at $\Rpeak = 1.19_{-0.18}^{+0.17}
  \REarth$.  \Will{More science stuff here.}
\end{abstract}

\bibliography{kepler-selection}

\begin{addendum}
\item WMF and IM are supported by a STFC consolidated grant number
  NNNN.  Computations in this work were performed on the University of
  Birmingham's BlueBEAR cluster.  Some/all of the data presented in
  this paper were obtained from the Mikulski Archive for Space
  Telescopes (MAST). STScI is operated by the Association of
  Universities for Research in Astronomy, Inc., under NASA contract
  NAS5-26555. Support for MAST for non-HST data is provided by the
  NASA Office of Space Science via grant NNX13AC07G and by other
  grants and contracts.  This paper includes data collected by the
  Kepler mission. Funding for the Kepler mission is provided by the
  NASA Science Mission directorate.
\item [Author Contributions] All authors assisted in the computational
  modelling, discussed the results, and edited the manuscript.
\item [Reprints] Reprints and permissions information is available at
  \url{www.nature.com/reprints}.
\item[Competing Interests] The authors declare that they have no
  competing financial interests.
\item[Correspondence] Correspondence and requests for materials should
  be addressed to W.M.F.\ (email: \email{w.farr@bham.ac.uk}).
\end{addendum}

\end{document}

### 2014-04-17 paper/kepler-selection.tex
Abstract changes, per Ilya's suggestions.

@@
   detecting Earth-like planets in Earth-like orbits about Sun-like
-  stars; previous estimates place $1\% \lesssim \etaearth \lesssim
-  34\%$, relying on extrapolations from regions in period-radius space
-  where detection is easier and the sample of observed planets is
-  complete\cite{2011ApJ...738...81W,2011ApJ...738..151C,2012ApJ...745...20T},
-  or modelling the selection effects carefully but using simpler
-  models of the period-radius distribution of
-  exoplanets\cite{2013ApJ...778...53D}, possibly using a low-noise
-  subset of the available data\cite{2013PNAS..11019273P}.
-  Additionally, no previous estimates account for contamination from
-  false-positive planetary identifications.  Here we present
+  stars; previous
+  estimates\cite{2011ApJ...738...81W,2011ApJ...738..151C,2012ApJ...745...20T,2013ApJ...778...53D,2013PNAS..11019273P}
+  place $1\% \lesssim \etaearth \lesssim 34\%$.  Here we present
   constraints on $\etaearth$ from a parameterised forward model of the
@@
   Kepler
-  satellite\cite{2010Sci...327..977B,2011ApJ...736...19B,2013ApJS..204...24B},
-  including an empirical parameterisation for the distribution of
-  false-positives in the data set.  We find $\etaearth =
+  satellite\cite{2010Sci...327..977B,2011ApJ...736...19B,2013ApJS..204...24B}.
+  We parameterise the intrinsic distribution of planetary periods and
+  radii using a single, correlated Gaussian component; treat selection
+  effects using a parameterised transit detection probability based on
+  the measured noise level and stellar properties in the Kepler
+  catalog; and include an empirically-parameterised, independent
+  component in the planet period-radius distribution to represent
+  false-positive planet detections.  We find $\etaearth =
   4.1^{+2.3}_{-1.7}\%$ (90\% CL).  Additionally, we find that each
   star hosts $4.04_{-0.68}^{+0.85}$ planets with $P \lesssim 3
-  \mathrm{yr}$ and $R \gtrsim 0.2 \REarth$, that the peak of the
-  planet-radius distribution lies at $\Rpeak = 1.19_{-0.18}^{+0.17}
-  \REarth$.  \Will{More science stuff here.}
+  \mathrm{yr}$ and $R \gtrsim 0.2 \REarth$ and that the peak of the
+  planet radius distribution lies at $\Rpeak = 1.19_{-0.18}^{+0.17}
+  \REarth$ (all 90\% CL).  Our empirical model for false-positive
+  contamination is consistent with the dominant source being backgroud
+  eclipsing binary stars, with $1.1_{-0.18}^{+0.2} \times 10^{-3}$
+  false-positives per star.  \Will{More science stuff here.}
 \end{abstract}

### 2014-04-18 paper/kepler-selection.tex
Better affiliations to correspond to Nature formats.
Better abstract, plots added to paper with (preliminary) captions, and code for plots changed.

@@
 \usepackage{xcolor}
+\usepackage{graphicx}
 
@@
 \newcommand{\REarth}{R_\oplus}
+\newcommand{\Rpl}{R_\mathrm{pl}}
 
@@
 
-\author{Will M. Farr$^{1,2}$, Chris Aldridge$^{1,3}$, Kirsty Stroud$^{1,4}$ \& Ilya Mandel$^{1,5}$}
+\author{Will M. Farr$^{1}$, Chris Aldridge$^{1}$, Kirsty Stroud$^{1}$ \& Ilya Mandel$^{1}$}
 
@@
 \item School of Physics and Astronomy, Univeristy of Birmingham, Birmingham, B15 2TT, United Kingdom
-\item \email{w.farr@bham.ac.uk}
-\item \email{CXA064@bham.ac.uk}
-\item \email{KLS081@bham.ac.uk}
-\item \email{ilyamandel@chgk.info}
 \end{affiliations}
@@
   The quantity $\etaearth$, the number density of planets per star per
-  logarithmic planetary radius per logarithmic orbital period,
-  describes the occurrence of Earth-like extrasolar planets.
-  Measurement of $\etaearth$ is complicated by the difficulty of
-  detecting Earth-like planets in Earth-like orbits about Sun-like
-  stars; previous
-  estimates\cite{2011ApJ...738...81W,2011ApJ...738..151C,2012ApJ...745...20T,2013ApJ...778...53D,2013PNAS..11019273P}
-  place $1\% \lesssim \etaearth \lesssim 34\%$.  Here we present
-  constraints on $\etaearth$ from a parameterised forward model of the
-  (correlated) period-radius distribution and the observational
-  selection function in the most recent (Q17) data release from the
-  Kepler
+  logarithmic planetary radius per logarithmic orbital period at one
+  Earth radius and one year periods, describes the occurrence of
+  Earth-like extrasolar planets.  Measurement of $\etaearth$ is
+  complicated by the difficulty of detecting Earth-like planets in
+  Earth-like orbits about Sun-like stars.  Previous
+  estimates\cite{2011ApJ...738..151C,2012ApJ...745...20T,2013ApJ...778...53D,2013PNAS..11019273P}
+  place $1\% \lesssim \etaearth \lesssim 34\%$.  These works dealt
+  with the problem of selection effects in the sample by either
+  analyzing a region of the period-radius parameter space where
+  observations are complete and extrapolating to $R = \REarth$ and $P
+  = 1 \mathrm{yr}$\cite{2011ApJ...738..151C,2012ApJ...745...20T} or
+  applying a binned analysis in the period-radius
+  plane\cite{2013ApJ...778...53D,2013PNAS..11019273P}.  Here we
+  present constraints on $\etaearth$ from a parameterised forward
+  model of the (correlated) period-radius distribution and the
+  observational selection function in the most recent (Q17) data
+  release from the Kepler
   satellite\cite{2010Sci...327..977B,2011ApJ...736...19B,2013ApJS..204...24B}.
-  We parameterise the intrinsic distribution of planetary periods and
-  radii using a single, correlated Gaussian component; treat selection
+  Our data set comprises 181,568 systems observed under the Kepler
+  exoplanet observing program, producing 2598 planetary candidates.
+  We parameterise the distribution of planetary periods and radii
+  using a single, correlated Gaussian component; treat selection
   effects using a parameterised transit detection probability based on
@@
   component in the planet period-radius distribution to represent
-  false-positive planet detections.  We find $\etaearth =
-  4.1^{+2.3}_{-1.7}\%$ (90\% CL).  Additionally, we find that each
-  star hosts $4.04_{-0.68}^{+0.85}$ planets with $P \lesssim 3
-  \mathrm{yr}$ and $R \gtrsim 0.2 \REarth$ and that the peak of the
-  planet radius distribution lies at $\Rpeak = 1.19_{-0.18}^{+0.17}
-  \REarth$ (all 90\% CL).  Our empirical model for false-positive
-  contamination is consistent with the dominant source being backgroud
-  eclipsing binary stars, with $1.1_{-0.18}^{+0.2} \times 10^{-3}$
-  false-positives per star.  \Will{More science stuff here.}
+  false-positive planet detections.  Using our model we can
+  simultaneously estimate $\etaearth$, place constraints on the planet
+  period-radius distribution function, and determine the degree of
+  contamination by false-positive candidate indentifications.  We find
+  $\etaearth = 3.8^{+2.1}_{-1.6}\%$ (90\% CL).  Additionally, we find
+  that each star hosts $3.82_{-0.60}^{+0.70}$ planets with $P \lesssim
+  3 \mathrm{yr}$ and $R \gtrsim 0.2 \REarth$, that the peak of the
+  planet radius distribution lies at $\Rpeak = 1.25_{-0.17}^{+0.16}
+  \REarth$, and that $\ln P$ and $\ln R$ are correlated with
+  correlation coefficient $r = 0.335_{-0.054}^{+0.052}$ (all 90\% CL).
+  Our empirical model for false-positive contamination is consistent
+  with the dominant source being backgroud eclipsing binary
+  stars\cite{2013ApJ...766...81F}, with $7.8_{-1.3}^{+1.4}\%$ (90\%
+  CL) of the candidates being false-positives.
 \end{abstract}
+
+\begin{figure}
+  \includegraphics[width=\columnwidth]{eta-earth}
+  \caption{\label{fig:eta-earth} \textbf{The posterior on $\etaearth$
+      accounting for selection effects and false positives.}  Recall
+    that $\etaearth \equiv \left. \frac{dN_\mathrm{pl}}{d\ln P \,d \ln
+      R} \right|_{P = 1\mathrm{yr}, R=1\REarth}$.  Vertical lines
+    indicate the 90\% credible range.  We find $\etaearth =
+    3.8^{+2.1}_{-1.6}\%$, slightly lower than but consistent with
+    previous work\cite{2013PNAS..11019273P}.}
+\end{figure}
+
+\begin{figure}
+  \includegraphics[width=\columnwidth]{npl}
+  \caption{\label{fig:npl} \textbf{The posterior on $\Rpl$.}  Recall
+    that $\Rpl$ is the number of planets per star.  Vertical lines
+    indicate the 90\% credible range.  We find $\Rpl =
+    3.82_{-0.60}^{+0.70}$. \Will{CITE/discuss previous work.}}
+\end{figure}
+
+\begin{figure}
+  \includegraphics[width=\columnwidth]{foreground-dist}
+  \caption{\label{fig:foreground-dist} \textbf{The inferred planet
+      period-radius distribution accounting for selection effects and
+      false-positives.}  (Upper Left) The planet number density per
+    logarithmic planet radius.  The density peaks at $\Rpeak =
+    1.19_{-0.18}^{+0.17} \REarth$ (90\% CL).  (Upper Right) The planet
+    number density in the period-radius plane.  The inferred
+    correlation coefficient between $\ln P$ and $\ln R$ is $r =
+    0.335_{-0.054}^{+0.052}$.  (Lower Left) Scatter plot of the radius
+    and period of the Kepler planet candidates.  Color indicates the
+    posterior false-positive probability for each candidate.  Overall,
+    the model prefers a false-positive rate of $7.8_{-1.3}^{+1.4}\%$
+    (90\% CL).  The primary contaminant is probably background
+    eclipsing binaries; our rate is consistent with previous
+    work\cite{2013ApJ...766...81F}. (Lower Right) The planet number
+    density per logarithmic planet period.  The density peaks at $P =
+    0.075_{-0.006}^{+0.007}\mathrm{yr}$.}
+\end{figure}
+
+\begin{figure}
+  \includegraphics[width=\columnwidth]{selection}
+  \caption{\label{fig:selection} \textbf{Comparison of synthetic data
+      sets produced from the forward model incorporating selection
+      effects with observed candidates.}  (Upper Left) The observed
+    (black curve) and synthetic (blue curve; including planets and
+    false positives, and using the fitted selection model to
+    down-select the candidates from the planet distribution)
+    normalised candidate density per logarithmic radius.  Except for a
+    discrepancy at $R \simeq 10 \REarth$---associated with hot
+    Jupiters, a distinct planetary population\Will{CITE, and
+      check}---the model produces a good fit to the observed
+    candidates over the range of reported radii.  (Upper Right)
+    Scatter plot of the observed candidates (black circles) and a
+    posterior-averaged draw of candidates from the model (blue
+    circles).  (Lower Left) Scatter plot of the observed candidates.
+    Colors indicate the posterior-averaged selection probability for
+    each planet about its host star.  Recall that the selection
+    probability is treated a product of a geometric factor giving the
+    probability of an isotropically-oriented orbit producing a transit
+    and a signal-to-noise-ratio-dependent transit detection
+    probability.  (Lower Right) The observed (black curve) and
+    synthetic (blue curve; including planets and false positives, and
+    using the fitted selection model to down-select the candidates
+    from the planet distribution) normalised candidate density per
+    logarithmic period.  Except for the aforementioned hot Jupiter
+    peak at $P \simeq 1 \mathrm{day}$ the model produces a good fit to
+    the observed candidates over the range of reported periods.}
+\end{figure}
 

### 2014-04-18 code/plots.py

@@ -26,6 +26,11 @@ def setup():
 
     pp.rcParams.update(params)
 
+def pcolormesh(x, y, z, *args, **kwargs):
+    z = 0.25*(z[:-1,:-1] + z[:-1, 1:] + z[1:,:-1] + z[1:,1:])
+
+    pp.pcolormesh(x, y, z, *args, **kwargs)
+
 def plot_error_ellipse(logpost, p):
     """Plots the location of the mean and the 1-sigma error ellipse in the
     P-R plane for the given params.
@@ -89,11 +94,11 @@ def plot_true_foreground_density(logpost, p):
                          np.exp(np.linspace(np.log(ymin), np.log(ymax), 250)))
     ZS = logpost.foreground_density(p, XS.flatten(), YS.flatten()).reshape((250, 250))
 
-    pp.pcolormesh(XS, YS, ZS)
+    pcolormesh(XS, YS, ZS)
     pp.xscale('log')
     pp.yscale('log')
 
-def compute_range(xs):
+def range(xs):
     """Returns ``(mean(xs), mean(xs)-x5, x95-mean(xs))`` where ``x5`` and
     ``x95`` are the 5-th and 95-th percentile of the ``xs``.
 
@@ -141,11 +146,7 @@ def correlation_coefficients(logpost, chain):
     return rs.reshape(chain.shape[:2])
 
 def plot_eta_earth_histogram(eta_earths, outdir=None):
-    tau = ac.autocorrelation_length_estimate(np.mean(eta_earths, axis=0))
-
-    thin_eta_earths = eta_earths[:, ::int(round(tau))]
-
-    pu.plot_histogram_posterior(thin_eta_earths.flatten(), normed=True, histtype='step', color='k')
+    pu.plot_histogram_posterior(eta_earths.flatten(), normed=True, histtype='step', color='k')
     pp.xlabel(r'$\eta_\oplus$')
     pp.ylabel(r'$p\left(\eta_\oplus\right)$')
 
@@ -156,19 +157,14 @@ def plot_eta_earth_histogram(eta_earths, outdir=None):
         pp.savefig(op.join(outdir, 'eta-earth.pdf'))
 
 def plot_foreground_distributions(logpost, chain, N=1000, outdir=None):
-    tau = np.max(ac.emcee_chain_autocorrelation_lengths(chain))
-
     fchain = chain.reshape((-1, chain.shape[2]))
-
[…]

### 2014-04-18 paper/kepler-selection.tex
Use automatically-produced X +/- Y quantities.

@@
 \newcommand{\apjs}{The Astrophysical Journal Supplement}
+
+\input{quantities}
 
@@
   = 1 \mathrm{yr}$\cite{2011ApJ...738..151C,2012ApJ...745...20T} or
-  applying a binned analysis in the period-radius
-  plane\cite{2013ApJ...778...53D,2013PNAS..11019273P}.  Here we
-  present constraints on $\etaearth$ from a parameterised forward
-  model of the (correlated) period-radius distribution and the
+  applying a binned analysis incorporating survey incompletness in the
+  period-radius plane\cite{2013ApJ...778...53D,2013PNAS..11019273P}.
+  Here we present constraints on $\etaearth$ from a parameterised
+  forward model of the (correlated) period-radius distribution and the
   observational selection function in the most recent (Q17) data
@@
   contamination by false-positive candidate indentifications.  We find
-  $\etaearth = 3.8^{+2.1}_{-1.6}\%$ (90\% CL).  Additionally, we find
-  that each star hosts $3.82_{-0.60}^{+0.70}$ planets with $P \lesssim
-  3 \mathrm{yr}$ and $R \gtrsim 0.2 \REarth$, that the peak of the
-  planet radius distribution lies at $\Rpeak = 1.25_{-0.17}^{+0.16}
-  \REarth$, and that $\ln P$ and $\ln R$ are correlated with
-  correlation coefficient $r = 0.335_{-0.054}^{+0.052}$ (all 90\% CL).
-  Our empirical model for false-positive contamination is consistent
-  with the dominant source being backgroud eclipsing binary
-  stars\cite{2013ApJ...766...81F}, with $7.8_{-1.3}^{+1.4}\%$ (90\%
+  $\etaearth = \earange$ (90\% CL), in agreement with
+  Ref.\ \cite{2013PNAS..11019273P}.  However, in contrast to
+  Ref.\ \cite{2013PNAS..11019273P}, we can conclude that each star
+  hosts $\rplrange$ planets with $P \lesssim 3 \mathrm{yr}$
+  and $R \gtrsim 0.2 \REarth$, that the peak of the planet radius
+  distribution lies at $\Rpeak = \rpeakrange \REarth$, and
+  that $\ln P$ and $\ln R$ are correlated with correlation coefficient
+  $r = \corrcoeffrange$ (all 90\% CL).  Our empirical model
+  for false-positive contamination is consistent with the dominant
+  source being backgroud eclipsing binary
+  stars\cite{2013ApJ...766...81F}, with $\fposrange$ (90\%
   CL) of the candidates being false-positives.
@@
       R} \right|_{P = 1\mathrm{yr}, R=1\REarth}$.  Vertical lines
-    indicate the 90\% credible range.  We find $\etaearth =
-    3.8^{+2.1}_{-1.6}\%$, slightly lower than but consistent with
-    previous work\cite{2013PNAS..11019273P}.}
+    indicate the 90\% credible range.  We find $\etaearth = \earange$,
+    in agreement with previous work\cite{2013PNAS..11019273P}.}
 \end{figure}
@@
     indicate the 90\% credible range.  We find $\Rpl =
-    3.82_{-0.60}^{+0.70}$. \Will{CITE/discuss previous work.}}
+    \rplrange$. \Will{CITE/discuss previous work.}}
 \end{figure}
@@
     logarithmic planet radius.  The density peaks at $\Rpeak =
-    1.19_{-0.18}^{+0.17} \REarth$ (90\% CL).  (Upper Right) The planet
+    \rpeakrange \REarth$ (90\% CL).  (Upper Right) The planet
     number density in the period-radius plane.  The inferred
     correlation coefficient between $\ln P$ and $\ln R$ is $r =
-    0.335_{-0.054}^{+0.052}$.  (Lower Left) Scatter plot of the radius
+    \corrcoeffrange$.  (Lower Left) Scatter plot of the radius
     and period of the Kepler planet candidates.  Color indicates the
     posterior false-positive probability for each candidate.  Overall,
-    the model prefers a false-positive rate of $7.8_{-1.3}^{+1.4}\%$
+    the model prefers a false-positive rate of $\fposrange$
     (90\% CL).  The primary contaminant is probably background
-    eclipsing binaries; our rate is consistent with previous
-    work\cite{2013ApJ...766...81F}. (Lower Right) The planet number
-    density per logarithmic planet period.  The density peaks at $P =
-    0.075_{-0.006}^{+0.007}\mathrm{yr}$.}
+    eclipsing binaries; our contamination rate is consistent with
+    previous work\cite{2013ApJ...766...81F}. (Lower Right) The planet
+    number density per logarithmic planet period.  The density peaks
+    at $P = \ppeakrange \mathrm{yr}$ (90\% CL).}
 \end{figure}
@@
     Scatter plot of the observed candidates (black circles) and a
-    posterior-averaged draw of candidates from the model (blue
-    circles).  (Lower Left) Scatter plot of the observed candidates.
-    Colors indicate the posterior-averaged selection probability for
-    each planet about its host star.  Recall that the selection
-    probability is treated a product of a geometric factor giving the
-    probability of an isotropically-oriented orbit producing a transit
-    and a signal-to-noise-ratio-dependent transit detection
-    probability.  (Lower Right) The observed (black curve) and
-    synthetic (blue curve; including planets and false positives, and
-    using the fitted selection model to down-select the candidates
+    posterior-averaged draw of observed candidates from the model
+    (blue circles).  (Lower Left) Scatter plot of the observed
+    candidates.  Colors indicate the posterior-averaged selection
+    probability for each planet about its host star.  Recall that the
+    selection probability is treated a product of a geometric factor
+    giving the probability of an isotropically-oriented orbit
+    producing a transit and a signal-to-noise-ratio-dependent transit
+    detection probability.  (Lower Right) The observed (black curve)
+    and synthetic (blue curve; including planets and false positives,
+    and using the fitted selection model to down-select the candidates
     from the planet distribution) normalised candidate density per

### 2014-04-18 code/plots.py

@@ -98,7 +98,7 @@ def plot_true_foreground_density(logpost, p):
     pp.xscale('log')
     pp.yscale('log')
 
-def range(xs):
+def par_range(xs):
     """Returns ``(mean(xs), mean(xs)-x5, x95-mean(xs))`` where ``x5`` and
     ``x95`` are the 5-th and 95-th percentile of the ``xs``.
 
@@ -343,3 +343,25 @@ def paper_plots(logpost, chain, eta_earths, outdir):
 
     pp.figure()
     plot_selection(logpost, chain, outdir=outdir)
+
+def paper_tex(logpost, chain, eta_earths, rs, outdir):
+    ea = par_range(100.0*eta_earths)
+    rpl = par_range(chain[:,:,0])
+    rpk = par_range(np.exp(chain[:,:,3]))
+    rc = par_range(rs)
+    fp = par_range(100.0*logpost.systems.shape[0]*chain[:,:,1]/logpost.candidates.shape[0])
+    pp = par_range(np.exp(chain[:,:,2]))
+
+    with open(op.join(outdir, 'quantities.tex'), 'w') as out:
+        out.write('\\newcommand{{\\earange}}{{{0:0.1f}_{{-{1:0.1f}}}^{{+{2:0.1f}}}\\%}}\n'.format(ea[0], ea[1], ea[2]))
+
+        out.write('\\newcommand{{\\rplrange}}{{{0:0.2f}_{{-{1:0.2f}}}^{{+{2:0.2f}}}}}\n'.format(rpl[0], rpl[1], rpl[2]))
+
+        out.write('\\newcommand{{\\rpeakrange}}{{{0:0.2f}_{{-{1:0.2f}}}^{{+{2:0.2f}}}}}\n'.format(rpk[0], rpk[1], rpk[2]))
+
+        out.write('\\newcommand{{\\corrcoeffrange}}{{{0:0.3f}_{{-{1:0.3f}}}^{{+{2:0.3f}}}}}\n'.format(rc[0], rc[1], rc[2]))
+
+        out.write('\\newcommand{{\\fposrange}}{{{0:0.1f}_{{-{1:0.1f}}}^{{+{2:0.1f}}}\\%}}\n'.format(fp[0], fp[1], fp[2]))
+
+        out.write('\\newcommand{{\\ppeakrange}}{{{0:0.3f}_{{-{1:0.3f}}}^{{+{2:0.3f}}}}}\n'.format(pp[0], pp[1], pp[2]))
+        

### 2014-11-18 paper/kepler-selection.tex
Monster commit of edits since forever ago; WMF thinks ready to submit.

@@
 \newcommand{\REarth}{R_\oplus}
+\newcommand{\RSun}{R_\odot}
+\newcommand{\RStar}{R_\mathrm{star}}
+\newcommand{\MStar}{M_\mathrm{star}}
+\newcommand{\MSun}{M_\odot}
 \newcommand{\Rpl}{R_\mathrm{pl}}
+\newcommand{\Npl}{N_\mathrm{pl}}
+
+\newcommand{\Rbg}{R_\mathrm{bg}}
+\newcommand{\Nbg}{N_\mathrm{bg}}
+
+\newcommand{\rhomin}{\rho_\mathrm{min}}
+\newcommand{\rhomax}{\rho_\mathrm{max}}
 
@@
 
+\newcommand{\aj}{The Astronomical Journal}
 \newcommand{\apj}{The Astrophysical Journal}
+\newcommand{\apjl}{The Astrophysical Journal Letters}
 \newcommand{\apjs}{The Astrophysical Journal Supplement}
+\newcommand{\mnras}{Monthly Notices of the Royal Astronomical Society}
+\newcommand{\pasp}{Publications of the Astronomical Society of the Pacific}
 
@@
 \begin{affiliations}
-\item School of Physics and Astronomy, Univeristy of Birmingham, Birmingham, B15 2TT, United Kingdom
+\item School of Physics and Astronomy, University of Birmingham, Birmingham, B15 2TT, United Kingdom
 \end{affiliations}
@@
   Earth-like orbits about Sun-like stars.  Previous
-  estimates\cite{2011ApJ...738..151C,2012ApJ...745...20T,2013ApJ...778...53D,2013PNAS..11019273P}
+  estimates\cite{Catanzarite2011,Traub2012,Dong2013,Petigura2013,Foreman-Mackey2014}
   place $1\% \lesssim \etaearth \lesssim 34\%$.  These works dealt
@@
   observations are complete and extrapolating to $R = \REarth$ and $P
-  = 1 \mathrm{yr}$\cite{2011ApJ...738..151C,2012ApJ...745...20T} or
-  applying a binned analysis incorporating survey incompletness in the
-  period-radius plane\cite{2013ApJ...778...53D,2013PNAS..11019273P}.
-  Here we present constraints on $\etaearth$ from a parameterised
-  forward model of the (correlated) period-radius distribution and the
-  observational selection function in the most recent (Q17) data
-  release from the Kepler
-  satellite\cite{2010Sci...327..977B,2011ApJ...736...19B,2013ApJS..204...24B}.
-  Our data set comprises 181,568 systems observed under the Kepler
-  exoplanet observing program, producing 2598 planetary candidates.
+  = 1 \mathrm{yr}$\cite{Catanzarite2011,Traub2012}, applying a binned
+  analysis incorporating survey incompleteness in the period-radius
+  plane\cite{Dong2013,Petigura2013} or analysing the results of a
+  customised planet detection pipeline on a subset of the Kepler
+  observations\cite{Petigura2013,Foreman-Mackey2014}.  Here we present
+  constraints on $\etaearth$ from a parameterised forward model of the
+  (correlated) period-radius distribution and the observational
+  selection function in the most recent (Q17) data release from the
+  Kepler satellite\cite{Borucki2010,Borucki2011,Batalha2013}.  Our
+  data set comprises 181,568 systems observed under the Kepler
+  exoplanet observing program (mostly G-type stars on the main
+  sequence\cite{Batalha2010}), producing 2598 planetary candidates.
   We parameterise the distribution of planetary periods and radii
@@
   catalog; and include an empirically-parameterised, independent
-  component in the planet period-radius distribution to represent
+  component in the period-radius distribution to represent
   false-positive planet detections.  Using our model we can
@@
   period-radius distribution function, and determine the degree of
-  contamination by false-positive candidate indentifications.  We find
-  $\etaearth = \earange$ (90\% CL), in agreement with
-  Ref.\ \cite{2013PNAS..11019273P}.  However, in contrast to
-  Ref.\ \cite{2013PNAS..11019273P}, we can conclude that each star
-  hosts $\rplrange$ planets with $P \lesssim 3 \mathrm{yr}$
-  and $R \gtrsim 0.2 \REarth$, that the peak of the planet radius
-  distribution lies at $\Rpeak = \rpeakrange \REarth$, and
-  that $\ln P$ and $\ln R$ are correlated with correlation coefficient
-  $r = \corrcoeffrange$ (all 90\% CL).  Our empirical model
-  for false-positive contamination is consistent with the dominant
-  source being backgroud eclipsing binary
-  stars\cite{2013ApJ...766...81F}, with $\fposrange$ (90\%
-  CL) of the candidates being false-positives.
+  contamination by false-positive candidate identifications.  We find
+  $\etaearth = \earange$ (90\% CL), in contrast to
+  Ref.\ \cite{Petigura2013} but rough agreement with the analysis of
+  the same data set in Ref.\ \cite{Foreman-Mackey2014}.  Also in
+  contrast to Ref.\ \cite{Petigura2013}, we can conclude that each
+  star hosts $\rplrange$ planets with $P \lesssim 3 \mathrm{yr}$ and
+  $R \gtrsim 0.2 \REarth$, that the peak of the planet radius
+  distribution lies at $\Rpeak = \rpeakrange \REarth$, and that $\ln
+  P$ and $\ln R$ are correlated with correlation coefficient $r =
+  \corrcoeffrange$ (all 90\% CL).  Our empirical model for
+  false-positive contamination is consistent with the dominant source
+  being background eclipsing binary stars\cite{Fressin2013}, with
+  $\fposrange$ (90\% CL) of the candidates being false-positives.
 \end{abstract}
+
+The Kepler satellite detects planets by observing a decrement in the
+photometric intensity of a planet's host star as the planet transits
+between the telescope and the star.  The Q17 data release describes
+2598 ``candidate'' planetary transit signals identified by the Kepler
+team from observations of stars in the ``EX'' observing program (which
+are primarily G-type main-sequence stars similar to our own
+Sun\cite{Batalha2010}), giving the inferred planetary period and
+radius for each.  The fractional depth of a planetary transit signal
+depends only on the radii of the planet and its host star.  The signal
+to noise ratio of a series of transits about a particular star in the
+Kepler satellite scales with planetary period and radius
+as\cite{Chatterjee2012}
+\begin{equation}
+  \rho = \rho_0 \left( \frac{R}{\REarth} \right)^2 \left(
+  \frac{P}{1\,\mathrm{yr}} \right)^{-1/3},
+\end{equation}
+where $\rho_0$ is the signal to noise ratio of a Earth-radius planet
+in a one-year orbit about that star, which depends on the number of
+quarters of observation of that star, the stellar radius, and the
+intrinsic variability of the stellar intensity
+(CDPP\cite{Christiansen2012}).  In our analysis, we obtain these
+quantities from the Kepler Input Catalog\cite{Batalha2010,Brown2011}
+and the MAST Kepler
+archive\footnote{\url{http://archive.stsci.edu/kepler/}}.  To a good
+approximation (see Fig.\ \ref{fig:selection} below), the detectability
+of a series of planetary transits in the Kepler data set is a function
+of the signal to noise ratio of the series.  Because the detectability
+of planet transits depends on both period and radius, it is important
+to consider the joint (i.e.\ two-dimensional) distribution of these
+quantities in the data\cite{Tabachnik2002,Youdin2011}.  We model the
+detection probability as a function of signal to noise as
+\begin{equation}
+  \label{eq:pdetect}
+  p_\mathrm{detect} = \begin{cases}
+    0 & \rho < \rhomin \\
+    \frac{\log \rho - \log \rhomin}{\log \rhomax - \log \rhomin} &
+    \rhomin < \rho < \rhomax \\
+    1 & \rhomax < \rho
+  \end{cases};
+\end{equation}
+the detection probability is zero for signals with $\rho < \rhomin$,
+rises linearly in $\log \rho$ for signals with $\rhomin < \rho <
+\rhomax$, and is 100\% for signals with $\rhomax < \rho$.  $\rhomin$
+and $\rhomax$ are parameters of our model.  We find $\rhomin =
+\rhominrange$ and $\rhomax = \rhomaxrange$ (90\% CL), in rough
+agreement with Refs.\ \cite{Borucki2011,Batalha2013}.
+
+The probability of a planet's orbit to align with the line-of-sight to
+Earth and thereby produce a transit signal is
+\begin{equation}
+  \label{eq:ptransit}
+  p_\mathrm{transit} = 0.0016\, \frac{\RStar}{\RSun}
+  \left(\frac{\MStar}{\MSun}\right)^{-1/3} \left(\frac{P}{1\,\mathrm{yr}}\right)^{-2/3}.
+\end{equation}
+Putting Eq.\ \ref{eq:pdetect} and \ref{eq:ptransit} together, the
+probability that Kepler will detect a planet of radius $R$ orbiting
+its host star at period $P$ is 
+\begin{equation}
+  p_\mathrm{select} = p_\mathrm{transit} p_\mathrm{detect}.
+\end{equation}
+
+We model the intrinsic distribution of planets as log-normal in period
+and radius, so observed planets populate the candidate $P$-$R$ plane
+with number density
+\begin{equation}
+  \label{eq:foreground-rate}
+  \frac{dN_\mathrm{obs}}{d\ln P\, d\ln R} = \left[ \sum_\mathrm{stars}
+    p_\mathrm{select}(P, R) \right] \Rpl N\left[ \mu, \Sigma
+    \right]\left( \ln P, \ln R \right),
+\end{equation}
+where $\Rpl$, $\mu$, and $\Sigma$ are parameters of our model, with
+$\Rpl$ the average number of planets per star, $\mu = \left[ \mu_P,
+  \mu_R \right]$ the mean of $\ln P$ and $\ln R$, and $\Sigma = \left[
+  \left[ \Sigma_{PP}, \Sigma_{PR} \right], \left[ \Sigma_{PR},
+    \Sigma_{RR} \right]\right]$ the covariance matrix of $\ln P$ and
+$\ln R$; $N\left[ \mu, \Sigma \right](x,y)$ is the normal distribution 
+\begin{equation}
+  N\left[ \mu, \Sigma \right](x,y) = \frac{1}{2 \pi \sqrt{\Sigma_{xx}
+      \Sigma_{yy} - \Sigma_{xy}^2}} \exp\left( - \frac{1}{2} \left(\left[
+    x,y \right] - \mu \right) \cdot \Sigma^{-1} \cdot \left(\left[ x,
+    y\right] - \mu \right)^T \right).
+\end{equation}
+The expected number of observed planets in our model is
+\begin{equation}
+  \Npl = \Rpl \int dP\,dR\, \sum_\mathrm{stars} p_\mathrm{select}(P,
+  R).
+\end{equation}
+Our model assumes that planets appear around their host stars in a
+Poisson process; this is almost certainly wrong in
+detail\cite{Weissbein2012}, but nevertheless provides a good fit to
+the observed data (see Figure \ref{fig:selection}).
+
+In addition to true planetary signals, we model a false-positive
+background of planet candidates empirically, assuming they populate
+the candidate $P$-$R$ plane with number density
+\begin{equation}
+  \label{eq:background-rate}
+  \frac{d\Nbg}{d \ln P \, d \ln R} = \frac{\Rbg}{\Delta \ln P \Delta
+    \ln R} \left( 1 + \vec{\gamma} \cdot \left[ \ln P - \ln P_\mathrm{mid} , \ln
+    R - \ln R_\mathrm{mid} \right] \right),
+\end{equation}
+where $\Delta \ln P = \ln P_\mathrm{max} - \ln P_\mathrm{min}$, $\ln
+P_\mathrm{mid} = 1/2\left(\ln P_\mathrm{max} - \ln P_\mathrm{min}
+\right)$, $\Delta \ln R = \ln R_\mathrm{max} - \ln R_\mathrm{min}$,
+$\ln R_\mathrm{mid} = 1/2\left(\ln R_\mathrm{max} - \ln R_\mathrm{min}
+\right)$.  $\Rbg$, the expected number of background false-positive
+events; $P_\mathrm{max}$, $P_\mathrm{min}$, $R_\mathrm{max}$, and
+$R_\mathrm{min}$, the boundaries in the $P$-$R$ plane within which
+background events appear; and $\gamma$, the gradient in the number
+density of background events, are parameters of our model.
+
+Unlike Ref.\ \cite{Foreman-Mackey2014}, we do not attempt to model the
+observational uncertainties in the estimated periods and radii from
+the Kepler candidate data set.  In spite of several candidates with
+very large uncertainties in measured parameters, we have found that our
+fit is essentially unchanged when applied to synthetic observations
+with periods and radii drawn re-drawn from the range of observational
+uncertainties quoted in the Q17 data release.
+
+The likelihood of the observed periods and radii under our model is an
+inhomogeneous Poisson likelihood\cite{Farr2013,Youdin2011} with a rate
+that is the sum of Eq.\ \eqref{eq:foreground-rate} and
+Eq.\ \eqref{eq:background-rate}.  We impose priors on our 15 model
+parameters as follows: for the planet occurrence rate $\Rpl$ and
+(implicitly) the parameters describing selection effects, we impose a
+$\frac{1}{\sqrt{\Npl}}$ prior; for the background rate $\Rbg$ we
+impose a $\frac{1}{\sqrt{\Rbg}}$ prior; for the selection model
+parameters $\rhomin$ and $\rhomax$ we impose a log-normal prior with
+unit width at SNRs of 3 and 11, respectively; in all other parameters
+we impose a flat (i.e.\ constant density) prior.  The product of
+likelihood and prior gives a Bayesian posterior density function on
+the fifteen-dimensional parameter space of our model.  We sample from
+this function using the \texttt{emcee} sampler of
+Ref.\ \cite{Foreman-Mackey2013}.  The posterior describes
+simultaneously the intrinsic distribution and number of exoplanets,
+the amount and distribution of the contaminating false-positive events
+in the candidate data set, and the selection function of the
+instrument for true planetary transit events.
+
+The main result of this paper, the posterior distribution for
+$\etaearth$, the number density of Earth-like planets, marginalised
+over all other parameters in our model (i.e.\ incorporating our
+uncertainty about contamination, selection effects, intrinsic
+distribution of planets, etc) appears in Fig.\ \ref{fig:eta-earth}.
+Recall that
+\begin{equation}
+  \etaearth = \left. \frac{dN}{d \ln P \ln R} \right|_{R = \REarth, P
+    = 1\,\mathrm{yr}} = \Rpl N\left[ \mu, \Sigma \right]\left( \ln 1\,\mathrm{yr},
+  \ln \REarth \right),
+\end{equation}
+which is roughly the number of planets per star with periods and radii
+within a factor of $\sqrt{e}$ of Earth's.  We find $\etaearth =
+\earange$ (90\% CL).  Our model also gives an estimate of the number
+of planets of any radius and period per star; the posterior for this
+quantity, marginalised over all other parameters appears in
+Fig.\ \ref{fig:npl}.  We find $\Rpl = \rplrange$ (90\% CL).
+
+Our model also allows us to produce a posterior on the distribution of
+planets in the period-radius plane, and a posterior for each observed
+planetary candidate of the probability that this candidate is a planet
+instead of a background contaminant; these posteriors appear in
+Fig. \ref{fig:foreground-dist}.  Our model finds that the
+false-positive rate in the candidate data set is $\fposrange$ (90\%
+CL), consistent with previous work\cite{Fressin2013} estimating the
+contamination in the Kepler candidate set.  Our model has the peak of
+the planet period-radius distribution at $\Rpeak = \rpeakrange
+\REarth$, $P_\mathrm{peak} = \ppeakrange \mathrm{yr}$, and the
+distribution of planetary radii and periods is correlated, with
+correlation coefficient $r = \corrcoeffrange$ (all at 90\% CL).
+
+Our model predicts a distribution for future observed data consistent
+with the already-observed candidate set.  These predictions can be
+used to perform graphical and posterior-predictive model
+checking\cite{Gelman2013}.  Fig.\ \ref{fig:selection} compares the
+predictions of our model for observed periods and radii (incorporating
+both planetary transits and background events) with the candidate set.
+This is a particularly stringent test of our parameterised selection
+model since the observed periods and radii are strongly influenced by
+the selection function of the Kepler telescope and pipeline.  Except
+for the known sub-population of hot
+Jupiters\cite{Albrecht2012,Naoz2012}, our model provides a very good
+fit to the observed data.  That a simple log-normal distribution in
+period and radius fits the observed distribution of planets well may
+indicate that planet formation is a stochastic process with many
+small, correlated, and multiplicative influences on planet period and
+radius resulting, from the central limit theorem, in a log-normal
+distribution in these parameters.
+
+The methods and analysed data sets of
+Refs.\ \cite{Petigura2013,Foreman-Mackey2014} are most comparable to
+ours.  These studies used the same data set,
+produced\cite{Petigura2013} from a subset of the available Kepler data
+and a customised pipeline to search for transit signals.  They both
+accounted for selection effects by measuring the recoverability of
+synthetic transit signals injected into their data, in contrast to our
+approach of empirically determining them from the observed data.
+Neither study attempted to account for contamination from
+falsely-identified candidate transit events, controlling this instead
+through careful choice of threshold.  Both studies used a more
+flexible model for the intrinsic distribution of planets than ours.
+Our result for $\etaearth$ is consistent with, but more precise than,
+Ref.\ \cite{Foreman-Mackey2014} and (somewhat) inconsistent with
+Ref.\ \cite{Petigura2013}.  The reasons for the difference between
+these two analyses of the same data set are not clear, but our
+agreement with Ref.\ \cite{Foreman-Mackey2014} lends support to that
+analysis and also partially validates our empirical selection model by
+comparison to the measured selection function in that work.
 
@@
     indicate the 90\% credible range.  We find $\etaearth = \earange$,
-    in agreement with previous work\cite{2013PNAS..11019273P}.}
+    in agreement with some previous work\cite{Foreman-Mackey2014}, and
+    in contrast to Ref.\ \cite{Petigura2013}.}
 \end{figure}
@@
     indicate the 90\% credible range.  We find $\Rpl =
-    \rplrange$. \Will{CITE/discuss previous work.}}
+    \rplrange$.}
 \end{figure}
@@
     eclipsing binaries; our contamination rate is consistent with
-    previous work\cite{2013ApJ...766...81F}. (Lower Right) The planet
+    previous work\cite{Fressin2013}. (Lower Right) The planet
     number density per logarithmic planet period.  The density peaks
@@
     discrepancy at $R \simeq 10 \REarth$---associated with hot
-    Jupiters, a distinct planetary population\Will{CITE, and
-      check}---the model produces a good fit to the observed
-    candidates over the range of reported radii.  (Upper Right)
-    Scatter plot of the observed candidates (black circles) and a
-    posterior-averaged draw of observed candidates from the model
-    (blue circles).  (Lower Left) Scatter plot of the observed
-    candidates.  Colors indicate the posterior-averaged selection
-    probability for each planet about its host star.  Recall that the
-    selection probability is treated a product of a geometric factor
-    giving the probability of an isotropically-oriented orbit
+    Jupiters, a distinct planetary
+    population\cite{Albrecht2012,Naoz2012}---the model produces a good
+    fit to the observed candidates over the range of reported radii.
+    Note particularly the tail at large radii that comes from
+    background contaminants in both observed and synthetic data.
+    (Upper Right) Scatter plot of the observed candidates (black
+    circles) and a posterior-averaged draw of observed candidates from
+    the model (blue circles).  (Lower Left) Scatter plot of the
+    observed candidates.  Colors indicate the posterior-averaged
+    selection probability for each planet about its host star.  Recall
+    that the selection probability is treated a product of a geometric
+    factor giving the probability of an isotropically-oriented orbit
     producing a transit and a signal-to-noise-ratio-dependent transit
@@
 \begin{addendum}
-\item WMF and IM are supported by a STFC consolidated grant number
-  NNNN.  Computations in this work were performed on the University of
+\item The code implementing this analysis is available under an
+  open-source ``MIT'' license at
+  \url{https://github.com/farr/kepler-selection}.  This work was
+  supported by the Science and Technology Facilities Council.
+  Computations in this work were performed on the University of
   Birmingham's BlueBEAR cluster.  Some/all of the data presented in

### 2014-11-19 paper/kepler-selection.tex
Comments from your anal reader.

@@
 \newcommand{\Will}[1]{\textcolor{cyan}{Will: #1}}
+\newcommand{\ilya}[1]{\textcolor{red}{#1}}
 
@@
   logarithmic planetary radius per logarithmic orbital period at one
-  Earth radius and one year periods, describes the occurrence of
+  Earth radius and one year period, describes the occurrence of
   Earth-like extrasolar planets.  Measurement of $\etaearth$ is
@@
   observations\cite{Petigura2013,Foreman-Mackey2014}.  Here we present
-  constraints on $\etaearth$ from a parameterised forward model of the
+  a measurement of $\etaearth$ from a parameterised forward model of the
   (correlated) period-radius distribution and the observational
@@
   contamination by false-positive candidate identifications.  We find
-  $\etaearth = \earange$ (90\% CL), in contrast to
-  Ref.\ \cite{Petigura2013} but rough agreement with the analysis of
-  the same data set in Ref.\ \cite{Foreman-Mackey2014}.  Also in
-  contrast to Ref.\ \cite{Petigura2013}, we can conclude that each
+  $\etaearth = \earange$ (90\% CL).  We conclude that each
   star hosts $\rplrange$ planets with $P \lesssim 3 \mathrm{yr}$ and
@@
   being background eclipsing binary stars\cite{Fressin2013}, with
-  $\fposrange$ (90\% CL) of the candidates being false-positives.
+  $\fposrange$ (90\% CL) of the candidates being false-positives.  \ilya{[Present the physical interpretation of correlated log-normal distribution as a result.]}
 \end{abstract}
@@
 in a one-year orbit about that star, which depends on the number of
-quarters of observation of that star, the stellar radius, and the
-intrinsic variability of the stellar intensity
-(CDPP\cite{Christiansen2012}).  In our analysis, we obtain these
+quarters of observation of that star, the stellar radius \ilya{and mass}, and the
+intrinsic variability of the stellar intensity\cite{Christiansen2012}.  In our analysis, we obtain these
 quantities from the Kepler Input Catalog\cite{Batalha2010,Brown2011}
@@
 of planet transits depends on both period and radius, it is important
-to consider the joint (i.e.\ two-dimensional) distribution of these
+to consider the joint (i.e., two-dimensional) distribution of these
 quantities in the data\cite{Tabachnik2002,Youdin2011}.  We model the
@@
 rises linearly in $\log \rho$ for signals with $\rhomin < \rho <
-\rhomax$, and is 100\% for signals with $\rhomax < \rho$.  $\rhomin$
+\rhomax$, and is 100\% for signals with $\rhomax < \rho$.
+\ilya{[Explain model in words before providing equations; say why this model was chosen and how (in)sensitive you are to model choice -- because otherwise, it looks a bit weird that it's shallower-than-linear.]}  $\rhomin$
 and $\rhomax$ are parameters of our model.  We find $\rhomin =
@@
 
-The probability of a planet's orbit to align with the line-of-sight to
+The probability that a planet's orbital plane will align with the line-of-sight to
 Earth and thereby produce a transit signal is
@@
 
+\ilya{[If introduced in abstract, say a log normal distribution is an excellent fit to the data instead of "we model".
+A log normal distribution in period ... copy from later, but add physics to mathematics.  (Why small multiplicative influences?)] } 
 We model the intrinsic distribution of planets as log-normal in period
@@
 where $\Rpl$, $\mu$, and $\Sigma$ are parameters of our model, with
-$\Rpl$ the average number of planets per star, $\mu = \left[ \mu_P,
+$\Rpl$ \ilya{[Different symbol? R overloaded.]} the average number of planets per star, $\mu = \left[ \mu_P,
   \mu_R \right]$ the mean of $\ln P$ and $\ln R$, and $\Sigma = \left[
@@
     \Sigma_{RR} \right]\right]$ the covariance matrix of $\ln P$ and
-$\ln R$; $N\left[ \mu, \Sigma \right](x,y)$ is the normal distribution 
+$\ln R$; $N\left[ \mu, \Sigma \right](x,y)$ is the normal distribution  \ilya{[Skip next equation.]}
 \begin{equation}
@@
 \begin{equation}
-  \Npl = \Rpl \int dP\,dR\, \sum_\mathrm{stars} p_\mathrm{select}(P,
+  N_\mathrm{obs} = \Rpl \int dP\,dR\, \sum_\mathrm{stars} p_\mathrm{select}(P,
   R).
 \end{equation}
+\ilya{[Maybe remove preceding equation.]}
 Our model assumes that planets appear around their host stars in a
-Poisson process; this is almost certainly wrong in
+Poisson process \ilya{[worse -- periods correlated, can't have two planets around same star with very similar periods]}; this is almost certainly wrong in
 detail\cite{Weissbein2012}, but nevertheless provides a good fit to
@@
 background of planet candidates empirically, assuming they populate
-the candidate $P$-$R$ plane with number density
+the candidate $P$-$R$ plane with number density \ilya{[Explain in words {\it before} equation.]}
 \begin{equation}
   \label{eq:background-rate}
-  \frac{d\Nbg}{d \ln P \, d \ln R} = \frac{\Rbg}{\Delta \ln P \Delta
+  \frac{d\Nbg}{d \ln P \, d \ln R} = \frac{\Nbg}{\Delta \ln P \Delta
     \ln R} \left( 1 + \vec{\gamma} \cdot \left[ \ln P - \ln P_\mathrm{mid} , \ln
@@
 $\ln R_\mathrm{mid} = 1/2\left(\ln R_\mathrm{max} - \ln R_\mathrm{min}
-\right)$.  $\Rbg$, the expected number of background false-positive
+\right)$.  $\Nbg$, the expected number of background false-positive
 events; $P_\mathrm{max}$, $P_\mathrm{min}$, $R_\mathrm{max}$, and
@@
 fit is essentially unchanged when applied to synthetic observations
-with periods and radii drawn re-drawn from the range of observational
+with periods and radii re-drawn from the range of observational
 uncertainties quoted in the Q17 data release.
@@
 Eq.\ \eqref{eq:background-rate}.  We impose priors on our 15 model
-parameters as follows: for the planet occurrence rate $\Rpl$ and
+parameters as follows: for the planet occurrence rate $\Rpl$ \ilya{[Lamda]} and
 (implicitly) the parameters describing selection effects, we impose a
-$\frac{1}{\sqrt{\Npl}}$ prior; for the background rate $\Rbg$ we
+$\frac{1}{\sqrt{\Npl}}$ prior; \ilya{[Nobs; too much technical detail -- consider leaving some for appendix?]} for the background rate $\Rbg$ we
 impose a $\frac{1}{\sqrt{\Rbg}}$ prior; for the selection model
@@
 unit width at SNRs of 3 and 11, respectively; in all other parameters
-we impose a flat (i.e.\ constant density) prior.  The product of
+we impose a flat (i.e., constant-density) prior.  The product of
 likelihood and prior gives a Bayesian posterior density function on
 the fifteen-dimensional parameter space of our model.  We sample from
-this function using the \texttt{emcee} sampler of
-Ref.\ \cite{Foreman-Mackey2013}.  The posterior describes
+this function using the \texttt{emcee} sampler\cite{Foreman-Mackey2013}.  The posterior describes
 simultaneously the intrinsic distribution and number of exoplanets,
@@
 $\etaearth$, the number density of Earth-like planets, marginalised
-over all other parameters in our model (i.e.\ incorporating our
+over all other parameters in our model (i.e., incorporating our
 uncertainty about contamination, selection effects, intrinsic
@@
 
-Our model also allows us to produce a posterior on the distribution of
-planets in the period-radius plane, and a posterior for each observed
-planetary candidate of the probability that this candidate is a planet
+Our model also \ilya{[Not also -- this is what you use to evaluate previous Eq. :) ]} allows us to produce a posterior on the distribution of
+planets in the period-radius plane, and the probability that any given planetary candidate is a planet
 instead of a background contaminant; these posteriors appear in
@@
 radius resulting, from the central limit theorem, in a log-normal
-distribution in these parameters.
+distribution in these parameters. \ilya{[Possibly revise wording if moved elsewhere.]}
 
@@
 Ref.\ \cite{Foreman-Mackey2014} and (somewhat) inconsistent with
-Ref.\ \cite{Petigura2013}.  The reasons for the difference between
-these two analyses of the same data set are not clear, but our
-agreement with Ref.\ \cite{Foreman-Mackey2014} lends support to that
-analysis and also partially validates our empirical selection model by
-comparison to the measured selection function in that work.
+Ref.\ \cite{Petigura2013}.
 
@@
   \includegraphics[width=\columnwidth]{eta-earth}
-  \caption{\label{fig:eta-earth} \textbf{The posterior on $\etaearth$
-      accounting for selection effects and false positives.}  Recall
-    that $\etaearth \equiv \left. \frac{dN_\mathrm{pl}}{d\ln P \,d \ln
-      R} \right|_{P = 1\mathrm{yr}, R=1\REarth}$.  Vertical lines
-    indicate the 90\% credible range.  We find $\etaearth = \earange$,
-    in agreement with some previous work\cite{Foreman-Mackey2014}, and
-    in contrast to Ref.\ \cite{Petigura2013}.}
+  \caption{\label{fig:eta-earth} \textbf{The posterior on $\etaearth \equiv \left. \frac{dN_\mathrm{pl}}{d\ln P \,d \ln
+      R} \right|_{P = 1\mathrm{yr}, R=1\REarth}$
+      accounting for selection effects and false positives.}  Vertical lines
+    indicate the 90\% credible range.  We find $\etaearth = \earange$.}
 \end{figure}
@@
   \includegraphics[width=\columnwidth]{npl}
-  \caption{\label{fig:npl} \textbf{The posterior on $\Rpl$.}  Recall
-    that $\Rpl$ is the number of planets per star.  Vertical lines
+  \caption{\label{fig:npl} \textbf{The posterior on $\Rpl$, the number of planets per star.}  Vertical lines
     indicate the 90\% credible range.  We find $\Rpl =
@@
   \caption{\label{fig:foreground-dist} \textbf{The inferred planet
-      period-radius distribution accounting for selection effects and
+      period--radius distribution accounting for selection effects and
       false-positives.}  (Upper Left) The planet number density per
@@
     \rpeakrange \REarth$ (90\% CL).  (Upper Right) The planet
-    number density in the period-radius plane.  The inferred
+    number density in the period--radius plane.  The inferred
     correlation coefficient between $\ln P$ and $\ln R$ is $r =
@@
     number density per logarithmic planet period.  The density peaks
-    at $P = \ppeakrange \mathrm{yr}$ (90\% CL).}
+    at $P = \ppeakrange \mathrm{yr}$ (90\% CL). \ilya{[Note: consider color map that doesn't have dark on both sides and light in the middle for B/W printing.]}}
 \end{figure}
@@
     observed candidates.  Colors indicate the posterior-averaged
-    selection probability for each planet about its host star.  Recall
-    that the selection probability is treated a product of a geometric
+    selection probability for each planet about its host star.  
+    The selection probability is treated a product of a geometric
     factor giving the probability of an isotropically-oriented orbit
@@
   Computations in this work were performed on the University of
-  Birmingham's BlueBEAR cluster.  Some/all of the data presented in
+  Birmingham's BlueBEAR cluster.  Some of the data presented in
   this paper were obtained from the Mikulski Archive for Space

### 2014-11-21 code/plots.py
Changed to cubehelix colormap for plots.

@@ -16,13 +16,14 @@ def setup():
 
     params = { 'backend' : 'pdf',
                'axes.labelsize' : 10,
-               'text.fontsize' : 10,
+               'font.size' : 10,
                'legend.fontsize' : 10,
                'xtick.labelsize' : 8,
                'ytick.labelsize' : 8,
                'text.usetex' : True,
                'figure.figsize' : fig_size,
-               'figure.autolayout' : True}
+               'figure.autolayout' : True,
+               'image.cmap' : 'cubehelix'}
 
     pp.rcParams.update(params)
 

### 2014-11-25 paper/kepler-selection.tex
New plots showing background density and detection efficiency, etc, plus captions

@@
 
-\author{Will M. Farr$^{1}$, Chris Aldridge$^{1}$, Kirsty Stroud$^{1}$ \& Ilya Mandel$^{1}$}
+\author{Will M. Farr$^{1}$, Ilya Mandel$^{1}$, Chris Aldridge$^{1}$ \& Kirsty Stroud$^{1}$}
 
@@
 \begin{figure}
-  \includegraphics[width=\columnwidth]{eta-earth}
-  \caption{\label{fig:eta-earth} \textbf{The posterior on $\etaearth \equiv \left. \frac{dN_\mathrm{pl}}{d\ln P \,d \ln
-      R} \right|_{P = 1\mathrm{yr}, R=1\REarth}$
-      accounting for selection effects and false positives.}  Vertical lines
-    indicate the 90\% credible range.  We find $\etaearth = \earange$.}
+  \includegraphics[width=\columnwidth]{pars}
+  \caption{\label{fig:eta-earth} \textbf{Posteriors on $\etaearth$ and
+      $Rpl$ accounting for selection effects and false-positive
+      detections.}  (Top) The inferred posterior density on $\etaearth
+    = dN/d\ln P \, d\ln R \left( 1\, \textnormal{yr}, R_\oplus
+    \right)$.  Vertical lines indicate the 90\% credible range.  We
+    find $\etaearth = \earange$.  (Bottom) The inferred posterior on
+    $\Rpl$, the number of planets per star with $P \lesssim 3
+    \mathrm{yr}$ and $R \gtrsim 0.2 \REarth$.  Vertical lines indicate
+    the 90\% credible range.  We find $\Rpl = \rplrange$.}
 \end{figure}
@@
 \begin{figure}
-  \includegraphics[width=\columnwidth]{npl}
-  \caption{\label{fig:npl} \textbf{The posterior on $\Rpl$, the number of planets per star.}  Vertical lines
-    indicate the 90\% credible range.  We find $\Rpl =
-    \rplrange$.}
+  \includegraphics[width=\columnwidth]{bg}
+  \caption{\label{fig:npl} \textbf{Inferred detection probability and
+      density of background contamination.} (Top) The inferred
+    detection probability versus signal-to-noise ratio (see
+    Eq.\ \eqref{eq:pdetect}) from our parameterised model of selection
+    effects.  The solid line is the posterior median detection
+    probability and the shading gives the 90\% credible posterior
+    interval.  Our inferred detection probability is in rough
+    agreement with the measurements of detection efficiency in
+    Refs.\ \cite{Borucki2011,Batalha2013}.  (Bottom) The number
+    density of false-positive candidate signals, $d\Nbg/d\ln P \, d\ln
+    R$ (see Eq.\ \eqref{eq:background-rate}).  The density is highest
+    at large candidate period and radius, consistent with the dominant
+    source of contamination being background eclipsing
+    binaries\cite{Fressin2013}.  Overall, our model finds $\fposrange$
+    of the candidates are false-positive background signals,
+    consistent with the analysis in Ref.\ \cite{Fressin2013}.}
 \end{figure}

### 2014-11-25 code/plots.py

@@ -332,12 +332,137 @@ def plot_nplanets_histogram(chain, outdir=None):
     if outdir is not None:
         pp.savefig(op.join(outdir, 'npl.pdf'))
 
+def _pselect(rhos, log_rho_min, log_rho_max):
+    ps = np.zeros(rhos.shape[0])
+
+    ps[rhos > np.exp(log_rho_max)] = 1.0
+
+    sel = (rhos > np.exp(log_rho_min)) & (rhos < np.exp(log_rho_max))
+    ps[sel] = (np.log(rhos[sel]) - log_rho_min)/(log_rho_max - log_rho_min)
+
+    return ps
+        
+def plot_pdetect(logpost, chain, outdir=None):
+    fchain = chain.reshape((-1, chain.shape[2]))
+    pfchain = logpost.to_params(fchain)
+
+    rho_max = 1.0 + np.exp(np.max(pfchain['log_snr_max']))
+    rhos = np.linspace(0, rho_max, 100)
+
+    psel = []
+    for p in fchain:
+        p = logpost.to_params(p)
+        psel.append(_pselect(rhos, p['log_snr_min'], p['log_snr_max']))
+    psel = np.array(psel)
+
+    med = np.median(psel, axis=0)
+    low = np.percentile(psel, 5, axis=0)
+    high = np.percentile(psel, 95, axis=0)
+
+    pp.fill_between(rhos, low, high, color='k', alpha=0.5)
+    pp.plot(rhos, med, '-k')
+    pp.xlabel(r'$\rho$')
+    pp.ylabel(r'$p_\mathrm{detect}$')
+
+    if outdir is not None:
+        pp.savefig(op.join(outdir, 'pdetect.pdf'))
+
+def plot_correlation_coefficient(logpost, chain, outdir=None):
+    fchain = chain.reshape((-1, chain.shape[2]))
+
+    rs = []
+    for p in fchain:
+        cm = logpost.covariance_matrix(p)
+        rs.append(cm[0,1] / np.sqrt(cm[0,0]*cm[1,1]))
+    rs = np.array(rs)
+
+    pu.plot_histogram_posterior(rs, normed=True, histtype='step', color='k')
+    pp.xlabel(r'$r$')
+    pp.ylabel(r'$p(r)$')
+
+    pp.axvline(np.percentile(rs, 5), color='k')
+    pp.axvline(np.percentile(rs, 95), color='k')
+
[…]

### 2014-11-25 paper/kepler-selection.tex
Final edits from Ilya's comments.

@@
 \newcommand{\MSun}{M_\odot}
-\newcommand{\Rpl}{R_\mathrm{pl}}
+\newcommand{\Rpl}{\Lambda_\mathrm{pl}}
 \newcommand{\Npl}{N_\mathrm{pl}}
 
-\newcommand{\Rbg}{R_\mathrm{bg}}
+\newcommand{\Rbg}{\Lambda_\mathrm{bg}}
 \newcommand{\Nbg}{N_\mathrm{bg}}
@@
 \newcommand{\aj}{The Astronomical Journal}
+\newcommand{\aap}{Astronomy \& Astrophysics}
 \newcommand{\apj}{The Astrophysical Journal}
@@
   being background eclipsing binary stars\cite{Fressin2013}, with
-  $\fposrange$ (90\% CL) of the candidates being false-positives.  \ilya{[Present the physical interpretation of correlated log-normal distribution as a result.]}
+  $\fposrange$ (90\% CL) of the candidates being false-positives.  The
+  distribution of planets we infer could be the result of a
+  highly-stochastic planet formation process producing many correlated,
+  fractional changes in planet sizes and orbits.  Our approach of
+  determining both the intrinsic distribution of objects and selection
+  effects empirically from survey data is generally applicable.
 \end{abstract}
@@
 in a one-year orbit about that star, which depends on the number of
-quarters of observation of that star, the stellar radius \ilya{and mass}, and the
-intrinsic variability of the stellar intensity\cite{Christiansen2012}.  In our analysis, we obtain these
+quarters of observation of that star, the stellar radius and mass, and
+the intrinsic variability of the stellar
+intensity\cite{Christiansen2012}.  In our analysis, we obtain these
 quantities from the Kepler Input Catalog\cite{Batalha2010,Brown2011}
@@
 quantities in the data\cite{Tabachnik2002,Youdin2011}.  We model the
-detection probability as a function of signal to noise as 
+detection probability of a transit as a function that rises linearly
+in the log of the signal to noise ratio from zero at a threshold
+signal to noise to one at a larger signal to noise:
 \begin{equation}
@@
     1 & \rhomax < \rho
-  \end{cases};
+  \end{cases},
 \end{equation}
-the detection probability is zero for signals with $\rho < \rhomin$,
-rises linearly in $\log \rho$ for signals with $\rhomin < \rho <
-\rhomax$, and is 100\% for signals with $\rhomax < \rho$.
-\ilya{[Explain model in words before providing equations; say why this model was chosen and how (in)sensitive you are to model choice -- because otherwise, it looks a bit weird that it's shallower-than-linear.]}  $\rhomin$
-and $\rhomax$ are parameters of our model.  We find $\rhomin =
-\rhominrange$ and $\rhomax = \rhomaxrange$ (90\% CL), in rough
-agreement with Refs.\ \cite{Borucki2011,Batalha2013}.
+where $\rhomin$ and $\rhomax$ are parameters of our model.  We find
+$\rhomin = \rhominrange$ and $\rhomax = \rhomaxrange$ (90\% CL), in
+rough agreement with Refs.\ \cite{Borucki2011,Batalha2013}.  A plot of
+our inferred detection probability appears in Fig.\ \ref{fig:det-bg}
 
@@
 
-\ilya{[If introduced in abstract, say a log normal distribution is an excellent fit to the data instead of "we model".
-A log normal distribution in period ... copy from later, but add physics to mathematics.  (Why small multiplicative influences?)] } 
-We model the intrinsic distribution of planets as log-normal in period
-and radius, so observed planets populate the candidate $P$-$R$ plane
-with number density
+A correlated log-normal distribution of planets in period and radius
+would be a natural outcome of a stochastic planet formation process
+that produced many correlated, fractional changes in planet sizes and
+orbits.  As we shall see (Figure\ \ref{fig:selection}), this simple
+model combined with the aforementioned selection function, provides a
+good fit to the Kepler candidate distribution.  In our model, observed
+planets populate the candidate $P$-$R$ plane with number density
 \begin{equation}
@@
 where $\Rpl$, $\mu$, and $\Sigma$ are parameters of our model, with
-$\Rpl$ \ilya{[Different symbol? R overloaded.]} the average number of planets per star, $\mu = \left[ \mu_P,
+$\Rpl$ the average number of planets per star, $\mu = \left[ \mu_P,
   \mu_R \right]$ the mean of $\ln P$ and $\ln R$, and $\Sigma = \left[
@@
     \Sigma_{RR} \right]\right]$ the covariance matrix of $\ln P$ and
-$\ln R$; $N\left[ \mu, \Sigma \right](x,y)$ is the normal distribution  \ilya{[Skip next equation.]}
-\begin{equation}
-  N\left[ \mu, \Sigma \right](x,y) = \frac{1}{2 \pi \sqrt{\Sigma_{xx}
-      \Sigma_{yy} - \Sigma_{xy}^2}} \exp\left( - \frac{1}{2} \left(\left[
-    x,y \right] - \mu \right) \cdot \Sigma^{-1} \cdot \left(\left[ x,
-    y\right] - \mu \right)^T \right).
-\end{equation}
-The expected number of observed planets in our model is
-\begin{equation}
-  N_\mathrm{obs} = \Rpl \int dP\,dR\, \sum_\mathrm{stars} p_\mathrm{select}(P,
-  R).
-\end{equation}
-\ilya{[Maybe remove preceding equation.]}
+$\ln R$; $N\left[ \mu, \Sigma \right](x,y)$ is the normal distribution
 Our model assumes that planets appear around their host stars in a
-Poisson process \ilya{[worse -- periods correlated, can't have two planets around same star with very similar periods]}; this is almost certainly wrong in
+Poisson process; this is almost certainly wrong in
 detail\cite{Weissbein2012}, but nevertheless provides a good fit to
@@
 background of planet candidates empirically, assuming they populate
-the candidate $P$-$R$ plane with number density \ilya{[Explain in words {\it before} equation.]}
+the candidate $P$-$R$ plane with a number density that has a linear
+gradient across a rectangular region in the $\ln P$-$\ln R$ plane:
 \begin{equation}
@@
 background events appear; and $\gamma$, the gradient in the number
-density of background events, are parameters of our model.
+density of background events, are parameters of our model.  This is a
+purely empirical model for the background contamination, but is
+reasonable if the chief contaminant is background eclipsing
+binaries\cite{Fressin2013,Duquennoy1991}.
 
@@
 Eq.\ \eqref{eq:background-rate}.  We impose priors on our 15 model
-parameters as follows: for the planet occurrence rate $\Rpl$ \ilya{[Lamda]} and
+parameters as follows: for the planet occurrence rate $\Rpl$ and
 (implicitly) the parameters describing selection effects, we impose a
-$\frac{1}{\sqrt{\Npl}}$ prior; \ilya{[Nobs; too much technical detail -- consider leaving some for appendix?]} for the background rate $\Rbg$ we
+$\frac{1}{\sqrt{\Npl}}$ prior; for the background rate $\Rbg$ we
 impose a $\frac{1}{\sqrt{\Rbg}}$ prior; for the selection model
 parameters $\rhomin$ and $\rhomax$ we impose a log-normal prior with
-unit width at SNRs of 3 and 11, respectively; in all other parameters
+unit width at signal to noise ratios  of 3 and 11, respectively; in all other parameters
 we impose a flat (i.e., constant-density) prior.  The product of
@@
 the fifteen-dimensional parameter space of our model.  We sample from
-this function using the \texttt{emcee} sampler\cite{Foreman-Mackey2013}.  The posterior describes
+this function using the \texttt{emcee}
+sampler\cite{Foreman-Mackey2013}.  The posterior describes
 simultaneously the intrinsic distribution and number of exoplanets,
@@
 of planets of any radius and period per star; the posterior for this
-quantity, marginalised over all other parameters appears in
-Fig.\ \ref{fig:npl}.  We find $\Rpl = \rplrange$ (90\% CL).
+quantity, marginalised over all other parameters also appears in
+Fig.\ \ref{fig:eta-earth}.  We find $\Rpl = \rplrange$ (90\% CL).
 
-Our model also \ilya{[Not also -- this is what you use to evaluate previous Eq. :) ]} allows us to produce a posterior on the distribution of
-planets in the period-radius plane, and the probability that any given planetary candidate is a planet
-instead of a background contaminant; these posteriors appear in
-Fig. \ref{fig:foreground-dist}.  Our model finds that the
-false-positive rate in the candidate data set is $\fposrange$ (90\%
-CL), consistent with previous work\cite{Fressin2013} estimating the
-contamination in the Kepler candidate set.  Our model has the peak of
-the planet period-radius distribution at $\Rpeak = \rpeakrange
-\REarth$, $P_\mathrm{peak} = \ppeakrange \mathrm{yr}$, and the
-distribution of planetary radii and periods is correlated, with
-correlation coefficient $r = \corrcoeffrange$ (all at 90\% CL).
+Our model allows us to produce a posterior on the distribution of
+planets in the period-radius plane, and the probability that any given
+planetary candidate is a planet instead of a background contaminant;
+these posteriors appear in Fig. \ref{fig:foreground-dist}.  Our model
+finds that the false-positive rate in the candidate data set is
+$\fposrange$ (90\% CL), consistent with previous
+work\cite{Fressin2013} estimating the contamination in the Kepler
+candidate set.  Our model has the peak of the planet period-radius
+distribution at $\Rpeak = \rpeakrange \REarth$, $P_\mathrm{peak} =
+\ppeakrange \mathrm{yr}$, and the distribution of planetary radii and
+periods is correlated, with correlation coefficient $r =
+\corrcoeffrange$ (all at 90\% CL).
 
@@
 radius resulting, from the central limit theorem, in a log-normal
-distribution in these parameters. \ilya{[Possibly revise wording if moved elsewhere.]}
+distribution in these parameters.
 
@@
   \includegraphics[width=\columnwidth]{bg}
-  \caption{\label{fig:npl} \textbf{Inferred detection probability and
-      density of background contamination.} (Top) The inferred
+  \caption{\label{fig:det-bg} \textbf{Inferred detection probability
+      and density of background contamination.} (Top) The inferred
     detection probability versus signal-to-noise ratio (see
@@
     logarithmic planet radius.  The density peaks at $\Rpeak =
-    \rpeakrange \REarth$ (90\% CL).  (Upper Right) The planet
-    number density in the period--radius plane.  The inferred
-    correlation coefficient between $\ln P$ and $\ln R$ is $r =
-    \corrcoeffrange$.  (Lower Left) Scatter plot of the radius
-    and period of the Kepler planet candidates.  Color indicates the
-    posterior false-positive probability for each candidate.  Overall,
-    the model prefers a false-positive rate of $\fposrange$
-    (90\% CL).  The primary contaminant is probably background
-    eclipsing binaries; our contamination rate is consistent with
-    previous work\cite{Fressin2013}. (Lower Right) The planet
-    number density per logarithmic planet period.  The density peaks
-    at $P = \ppeakrange \mathrm{yr}$ (90\% CL). \ilya{[Note: consider color map that doesn't have dark on both sides and light in the middle for B/W printing.]}}
+    \rpeakrange \REarth$ (90\% CL).  (Upper Right) The planet number
+    density in the period--radius plane.  The inferred correlation
+    coefficient between $\ln P$ and $\ln R$ is $r = \corrcoeffrange$.
+    (Lower Left) Scatter plot of the radius and period of the Kepler
+    planet candidates.  Color indicates the posterior false-positive
+    probability for each candidate.  Overall, the model prefers a
+    false-positive rate of $\fposrange$ (90\% CL).  The primary
+    contaminant is probably background eclipsing binaries; our
+    contamination rate is consistent with previous
+    work\cite{Fressin2013}. (Lower Right) The planet number density
+    per logarithmic planet period.  The density peaks at $P =
+    \ppeakrange \mathrm{yr}$ (90\% CL). }
 \end{figure}

### 2014-11-25 paper/cover-letter.tex
Final edits; as submitted.

\documentclass{letter}

\address{University of Birmingham\\School of Physics and
  Astronomy\\Birmingham\\B15 2TT\\United
  Kingdom\\\\\texttt{w.farr@bham.ac.uk}\\+44 783 115 3237}

\signature{Will Farr}

\begin{document}

\begin{letter}{Dr. Leslie Sage\\The Macmillan Building\\4 Crinan Street\\London N1 9XW\\United Kingdom}

  \opening{Dear Dr. Sage:}

  Please find enclosed a submission for Nature Letters entitled ``The
  Occurrence of Earth-Like Planets Around Other Stars.''  The
  submission comprises approximately 1900 words of text and four
  figures (with captions).  We do not at this time have any extended
  sections, supplementary information or supporting manuscripts.  The
  requested brief paragraphs justifying publication in Nature and
  summarising the letter's appeal to a broad audience follow.

  \textbf{For the editor:} The quantity $\eta_\oplus$ is of broad
  scientific interest.  Its estimation is a chief goal of NASA's
  Kepler mission, but is complicated by the difficulty of detecting
  Earth-like planets about other stars.  In addition to constraining
  $\eta_\oplus$ using the complete Kepler data set, we have modelled
  the distribution of planet orbital periods and radii and find that
  it is well-fit by a single Gaussian component.  Our approach to
  reconstructing the period-radius distribution of planets is
  generally applicable to any survey with significant selection
  effects.

  \textbf{For the public:} We use the latest results from NASA's
  Kepler mission to count how many planets of roughly Earth's size
  orbit stars like the Sun in our galaxy at about the same distance
  from their host stars as Earth.  We find that about 4\% of stars
  have a planet the size of Earth in an Earth-like orbit; with about
  100 billion Sun-like stars in the galaxy, this means there are over
  4 billion Earth-sized planets in one year orbits in our galaxy!

  We thank you for your consideration.

  \closing{Sincerely,}

\end{letter}

\end{document}

### 2014-11-25 paper/kepler-selection.tex
Final edits; as submitted.

@@
   with the problem of selection effects in the sample by either
-  analyzing a region of the period-radius parameter space where
+  analysing a region of the period-radius parameter space where
   observations are complete and extrapolating to $R = \REarth$ and $P

### 2014-11-26 paper/kepler-selection.tex
Changed abstract wording a bit (moved discussion of previous results to final paragraph).

@@
   complicated by the difficulty of detecting Earth-like planets in
-  Earth-like orbits about Sun-like stars.  Previous
-  estimates\cite{Catanzarite2011,Traub2012,Dong2013,Petigura2013,Foreman-Mackey2014}
-  place $1\% \lesssim \etaearth \lesssim 34\%$.  These works dealt
-  with the problem of selection effects in the sample by either
-  analysing a region of the period-radius parameter space where
-  observations are complete and extrapolating to $R = \REarth$ and $P
-  = 1 \mathrm{yr}$\cite{Catanzarite2011,Traub2012}, applying a binned
-  analysis incorporating survey incompleteness in the period-radius
-  plane\cite{Dong2013,Petigura2013} or analysing the results of a
-  customised planet detection pipeline on a subset of the Kepler
-  observations\cite{Petigura2013,Foreman-Mackey2014}.  Here we present
-  a measurement of $\etaearth$ from a parameterised forward model of the
-  (correlated) period-radius distribution and the observational
+  Earth-like orbits about Sun-like stars.  Here we present
+  a measurement of $\etaearth$ from a parameterised forward model of
+  the (correlated) period-radius distribution and the observational
   selection function in the most recent (Q17) data release from the
@@
   contamination by false-positive candidate identifications.  We find
-  $\etaearth = \earange$ (90\% CL).  We conclude that each
-  star hosts $\rplrange$ planets with $P \lesssim 3 \mathrm{yr}$ and
-  $R \gtrsim 0.2 \REarth$, that the peak of the planet radius
-  distribution lies at $\Rpeak = \rpeakrange \REarth$, and that $\ln
-  P$ and $\ln R$ are correlated with correlation coefficient $r =
-  \corrcoeffrange$ (all 90\% CL).  Our empirical model for
-  false-positive contamination is consistent with the dominant source
-  being background eclipsing binary stars\cite{Fressin2013}, with
-  $\fposrange$ (90\% CL) of the candidates being false-positives.  The
-  distribution of planets we infer could be the result of a
-  highly-stochastic planet formation process producing many correlated,
-  fractional changes in planet sizes and orbits.  Our approach of
-  determining both the intrinsic distribution of objects and selection
-  effects empirically from survey data is generally applicable.
+  $\etaearth = \earange$ (90\% CL).  We conclude that each star hosts
+  $\rplrange$ planets with $P \lesssim 3 \mathrm{yr}$ and $R \gtrsim
+  0.2 \REarth$, that the peak of the planet radius distribution lies
+  at $\Rpeak = \rpeakrange \REarth$, and that $\ln P$ and $\ln R$ are
+  correlated with correlation coefficient $r = \corrcoeffrange$ (all
+  90\% CL).  Our empirical model for false-positive contamination is
+  consistent with the dominant source being background eclipsing
+  binary stars\cite{Fressin2013}, with $\fposrange$ (90\% CL) of the
+  candidates being false-positives.  The distribution of planets we
+  infer is consistent with a highly-stochastic planet formation
+  process producing many correlated, fractional changes in planet
+  sizes and orbits.  Our approach of determining both the intrinsic
+  distribution of objects and selection effects empirically from
+  survey data is generally applicable.
 \end{abstract}
@@
 
-The methods and analysed data sets of
-Refs.\ \cite{Petigura2013,Foreman-Mackey2014} are most comparable to
-ours.  These studies used the same data set,
+Previous
+estimates\cite{Catanzarite2011,Traub2012,Dong2013,Petigura2013,Foreman-Mackey2014}
+place $1\% \lesssim \etaearth \lesssim 34\%$.  These works dealt with
+the problem of selection effects in the sample by either analysing a
+region of the period-radius parameter space where observations are
+complete and extrapolating to $R = \REarth$ and $P = 1
+\mathrm{yr}$\cite{Catanzarite2011,Traub2012}, applying a binned
+analysis incorporating survey incompleteness in the period-radius
+plane\cite{Dong2013,Petigura2013} or analysing the results of a
+customised planet detection pipeline on a subset of the Kepler
+observations\cite{Petigura2013,Foreman-Mackey2014}.  The methods and
+analysed data sets of Refs.\ \cite{Petigura2013,Foreman-Mackey2014}
+are most comparable to ours.  These studies used the same data set,
 produced\cite{Petigura2013} from a subset of the available Kepler data

### 2014-11-26 paper/kepler-selection.tex
Updates to Science format, following Nature rejection :(.

@@
-\documentclass{nature}
+\documentclass[12pt]{article}
+
+\usepackage{scicite}
+\usepackage{times}
+
+\topmargin 0.0cm
+\oddsidemargin 0.2cm
+\textwidth 16cm 
+\textheight 21cm
+\footskip 1.0cm
+
+\newenvironment{sciabstract}{%
+\begin{quote} \bf}
+{\end{quote}}
+
+\renewcommand\refname{References and Notes}
+
+\newcounter{lastnote}
+\newenvironment{scilastnote}{%
+\setcounter{lastnote}{\value{enumiv}}%
+\addtocounter{lastnote}{+1}%
+\begin{list}%
+{\arabic{lastnote}.}
+{\setlength{\leftmargin}{.22in}}
+{\setlength{\labelsep}{.5em}}}
+{\end{list}}
 
@@
 
-\bibliographystyle{naturemag}
+\bibliographystyle{Science}
 
@@
 
-\input{quantities}
+%\input{quantities}
+\newcommand{\earange}{3.9_{-1.6}^{+2.2}\%}
+\newcommand{\rplrange}{3.83_{-0.62}^{+0.76}}
+\newcommand{\rpeakrange}{1.25_{-0.17}^{+0.16}}
+\newcommand{\corrcoeffrange}{0.334_{-0.053}^{+0.052}}
+\newcommand{\fposrange}{7.8_{-1.3}^{+1.4}\%}
+\newcommand{\ppeakrange}{0.075_{-0.006}^{+0.007}}
+\newcommand{\rhominrange}{5.46^{+0.18}_{-0.18}}
+\newcommand{\rhomaxrange}{18.8^{+1.9}_{-1.9}}
+
+\title{The Occurrence of Earth-Like Planets Around Other Stars}
+\author{Will M. Farr$^{1\ast}$, Ilya Mandel$^{1}$, Chris Aldridge$^{1}$ \& Kirsty Stroud$^{1}$\\
+  \normalsize{$^1$School of Physics and Astronomy}\\
+  \normalsize{University of Birmingham}\\
+  \normalsize{Birmingham}\\
+  \normalsize{B15 2TT}\\
+  \normalsize{United Kingdom}\\
+  \normalsize{$^\ast$To whom correspondence should be addressed; E-mail: w.farr@bham.ac.uk.}}
+
+\date{}
 
 \begin{document}
-\title{The Occurrence of Earth-Like Planets Around Other Stars}
 
-\author{Will M. Farr$^{1}$, Ilya Mandel$^{1}$, Chris Aldridge$^{1}$ \& Kirsty Stroud$^{1}$}
+\baselineskip24pt
 
@@
 
-\begin{affiliations}
-\item School of Physics and Astronomy, University of Birmingham, Birmingham, B15 2TT, United Kingdom
-\end{affiliations}
+\begin{sciabstract}
+  The quantity $\etaearth$, the number density of planets per star per
+  logarithmic planetary radius per logarithmic orbital period at one
+  Earth radius and one year period, describes the occurrence of
+  Earth-like extrasolar planets.  Here we present a
+  measurement of $\etaearth$ from a parameterised forward model of the
+  (correlated) period-radius distribution and the observational
+  selection function in the most recent (Q17) data release from the
+  Kepler satellite\cite{Borucki2010,Borucki2011,Batalha2013}.  We find
+  $\etaearth = \earange$ (90\% CL).  We conclude that each star hosts
+  $\rplrange$ planets with $P \lesssim 3 \mathrm{yr}$ and $R \gtrsim
+  0.2 \REarth$.  Our empirical model for false-positive contamination
+  is consistent with the dominant source being background eclipsing
+  binary stars\cite{Fressin2013}.  The distribution of planets we
+  infer is consistent with a highly-stochastic planet formation
+  process producing many correlated, fractional changes in planet
+  sizes and orbits.  
+\end{sciabstract}
 
-\begin{abstract}
 The quantity $\etaearth$, the number density of planets per star per
@@
 complicated by the difficulty of detecting Earth-like planets in
-  Earth-like orbits about Sun-like stars.  Here we present
-  a measurement of $\etaearth$ from a parameterised forward model of
-  the (correlated) period-radius distribution and the observational
-  selection function in the most recent (Q17) data release from the
-  Kepler satellite\cite{Borucki2010,Borucki2011,Batalha2013}.  Our
-  data set comprises 181,568 systems observed under the Kepler
-  exoplanet observing program (mostly G-type stars on the main
-  sequence\cite{Batalha2010}), producing 2598 planetary candidates.
-  We parameterise the distribution of planetary periods and radii
-  using a single, correlated Gaussian component; treat selection
-  effects using a parameterised transit detection probability based on
-  the measured noise level and stellar properties in the Kepler
-  catalog; and include an empirically-parameterised, independent
-  component in the period-radius distribution to represent
-  false-positive planet detections.  Using our model we can
-  simultaneously estimate $\etaearth$, place constraints on the planet
-  period-radius distribution function, and determine the degree of
-  contamination by false-positive candidate identifications.  We find
-  $\etaearth = \earange$ (90\% CL).  We conclude that each star hosts
-  $\rplrange$ planets with $P \lesssim 3 \mathrm{yr}$ and $R \gtrsim
-  0.2 \REarth$, that the peak of the planet radius distribution lies
-  at $\Rpeak = \rpeakrange \REarth$, and that $\ln P$ and $\ln R$ are
-  correlated with correlation coefficient $r = \corrcoeffrange$ (all
-  90\% CL).  Our empirical model for false-positive contamination is
-  consistent with the dominant source being background eclipsing
-  binary stars\cite{Fressin2013}, with $\fposrange$ (90\% CL) of the
-  candidates being false-positives.  The distribution of planets we
-  infer is consistent with a highly-stochastic planet formation
-  process producing many correlated, fractional changes in planet
-  sizes and orbits.  Our approach of determining both the intrinsic
-  distribution of objects and selection effects empirically from
-  survey data is generally applicable.
-\end{abstract}
+Earth-like orbits about Sun-like stars.  Here we present a measurement
+of $\etaearth$ from a parameterised forward model of the (correlated)
+period-radius distribution and the observational selection function in
+the most recent (Q17) data release from the Kepler
+satellite\cite{Borucki2010,Borucki2011,Batalha2013}.  Our data set
+comprises 181,568 systems observed under the Kepler exoplanet
+observing program (mostly G-type stars on the main
+sequence\cite{Batalha2010}), producing 2598 planetary candidates.  We
+parameterise the distribution of planetary periods and radii using a
+single, correlated Gaussian component; treat selection effects using a
+parameterised transit detection probability based on the measured
+noise level and stellar properties in the Kepler catalog; and include
+an empirically-parameterised, independent component in the
+period-radius distribution to represent false-positive planet
+detections.  Using our model we can simultaneously estimate
+$\etaearth$, place constraints on the planet period-radius
+distribution function, and determine the degree of contamination by
+false-positive candidate identifications.  We find $\etaearth =
+\earange$ (90\% CL).  We conclude that each star hosts $\rplrange$
+planets with $P \lesssim 3 \mathrm{yr}$ and $R \gtrsim 0.2 \REarth$,
+that the peak of the planet radius distribution lies at $\Rpeak =
+\rpeakrange \REarth$, and that $\ln P$ and $\ln R$ are correlated with
+correlation coefficient $r = \corrcoeffrange$ (all 90\% CL).  Our
+empirical model for false-positive contamination is consistent with
+the dominant source being background eclipsing binary
+stars\cite{Fressin2013}, with $\fposrange$ (90\% CL) of the candidates
+being false-positives.  The distribution of planets we infer is
+consistent with a highly-stochastic planet formation process producing
+many correlated, fractional changes in planet sizes and orbits.  Our
+approach of determining both the intrinsic distribution of objects and
+selection effects empirically from survey data is generally
+applicable.
 
@@
 
+\begin{figure}
+  \includegraphics[width=\columnwidth]{bg}
+  \caption{\label{fig:det-bg} \textbf{Inferred detection probability
+      and density of background contamination.} (Top) The inferred
+    detection probability versus signal-to-noise ratio (see
+    Eq.\ \eqref{eq:pdetect}) from our parameterised model of selection
+    effects.  The solid line is the posterior median detection
+    probability and the shading gives the 90\% credible posterior
+    interval.  Our inferred detection probability is in rough
+    agreement with the measurements of detection efficiency in
+    Refs.\ \cite{Borucki2011,Batalha2013}.  (Bottom) The number
+    density of false-positive candidate signals, $d\Nbg/d\ln P \, d\ln
+    R$ (see Eq.\ \eqref{eq:background-rate}).  The density is highest
+    at large candidate period and radius, consistent with the dominant
+    source of contamination being background eclipsing
+    binaries\cite{Fressin2013}.  Overall, our model finds $\fposrange$
+    of the candidates are false-positive background signals,
+    consistent with the analysis in Ref.\ \cite{Fressin2013}.}
+\end{figure}
+
 The probability that a planet's orbital plane will align with the line-of-sight to
@@
 \begin{equation}
+  \label{eq:pselect}
   p_\mathrm{select} = p_\mathrm{transit} p_\mathrm{detect}.
@@
 reasonable if the chief contaminant is background eclipsing
-binaries\cite{Fressin2013,Duquennoy1991}.
+binaries\cite{Fressin2013,Duquennoy1991}.  The posterior on the
+background number density in the $P$-$R$ plane appears in
+Figure\ \ref{fig:det-bg}.
 
@@
 parameters $\rhomin$ and $\rhomax$ we impose a log-normal prior with
-unit width at signal to noise ratios  of 3 and 11, respectively; in all other parameters
-we impose a flat (i.e., constant-density) prior.  The product of
-likelihood and prior gives a Bayesian posterior density function on
-the fifteen-dimensional parameter space of our model.  We sample from
-this function using the \texttt{emcee}
+unit width at signal to noise ratios of 3 and 11, respectively; in all
+other parameters we impose a flat (i.e., constant-density) prior.  The
+product of likelihood and prior gives a Bayesian posterior density
+function on the fifteen-dimensional parameter space of our model.  We
+sample from this function using the \texttt{emcee}
 sampler\cite{Foreman-Mackey2013}.  The posterior describes
@@
 
+\begin{figure}
+  \includegraphics[width=\columnwidth]{pars}
+  \caption{\label{fig:eta-earth} \textbf{Posteriors on $\etaearth$ and
+      $Rpl$ accounting for selection effects and false-positive
+      detections.}  (Top) The inferred posterior density on $\etaearth
+    = dN/d\ln P \, d\ln R \left( 1\, \textnormal{yr}, R_\oplus
+    \right)$.  Vertical lines indicate the 90\% credible range.  We
+    find $\etaearth = \earange$.  (Bottom) The inferred posterior on
+    $\Rpl$, the number of planets per star with $P \lesssim 3
+    \mathrm{yr}$ and $R \gtrsim 0.2 \REarth$.  Vertical lines indicate
+    the 90\% credible range.  We find $\Rpl = \rplrange$.}
+\end{figure}
+
 Our model allows us to produce a posterior on the distribution of
@@
 \corrcoeffrange$ (all at 90\% CL).
+
+\begin{figure}
+  \includegraphics[width=\columnwidth]{foreground-dist}
+  \caption{\label{fig:foreground-dist} \textbf{The inferred planet
+      period--radius distribution accounting for selection effects and
+      false-positives.}  (Upper Left) The planet number density per
+    logarithmic planet radius.  The density peaks at $\Rpeak =
+    \rpeakrange \REarth$ (90\% CL).  (Upper Right) The planet number
+    density in the period--radius plane.  The inferred correlation
+    coefficient between $\ln P$ and $\ln R$ is $r = \corrcoeffrange$.
+    (Lower Left) Scatter plot of the radius and period of the Kepler
+    planet candidates.  Color indicates the posterior false-positive
+    probability for each candidate.  Overall, the model prefers a
+    false-positive rate of $\fposrange$ (90\% CL).  The primary
+    contaminant is probably background eclipsing binaries; our
+    contamination rate is consistent with previous
+    work\cite{Fressin2013}. (Lower Right) The planet number density
+    per logarithmic planet period.  The density peaks at $P =
+    \ppeakrange \mathrm{yr}$ (90\% CL). }
+\end{figure}
 
@@
 distribution in these parameters.
+
+\begin{figure}
+  \includegraphics[width=\columnwidth]{selection}
+  \caption{\label{fig:selection} \textbf{Comparison of synthetic data
+      sets produced from the forward model incorporating selection
+      effects with observed candidates.}  (Upper Left) The observed
+    (black curve) and synthetic (blue curve) normalised candidate
+    density per logarithmic radius.  Except for a discrepancy at $R
+    \simeq 10 \REarth$---associated with hot Jupiters, a distinct
+    planetary population\cite{Albrecht2012,Naoz2012}---the model
+    produces a good fit to the observed candidates over the range of
+    reported radii.  Note particularly the tail at large radii that
+    comes from background contaminants in both observed and synthetic
+    data.  (Upper Right) Scatter plot of the observed candidates
+    (black circles) and a posterior-averaged draw of observed
+    candidates from the model (blue circles).  (Lower Left) Scatter
+    plot of the observed candidates.  Colors indicate the
+    posterior-averaged selection probability for each planet about its
+    host star (see Eq.\ \eqref{eq:pselect}).  (Lower Right) The
+    observed (black curve) and synthetic (blue curve) normalised
+    candidate density per logarithmic period.  Except for the
+    aforementioned hot Jupiter peak at $P \simeq 1 \mathrm{day}$ the
+    model produces a good fit to the observed candidates over the
+    range of reported periods.}
+\end{figure}
 
@@
 
-\begin{figure}
-  \includegraphics[width=\columnwidth]{pars}
-  \caption{\label{fig:eta-earth} \textbf{Posteriors on $\etaearth$ and
-      $Rpl$ accounting for selection effects and false-positive
-      detections.}  (Top) The inferred posterior density on $\etaearth
-    = dN/d\ln P \, d\ln R \left( 1\, \textnormal{yr}, R_\oplus
-    \right)$.  Vertical lines indicate the 90\% credible range.  We
-    find $\etaearth = \earange$.  (Bottom) The inferred posterior on
-    $\Rpl$, the number of planets per star with $P \lesssim 3
-    \mathrm{yr}$ and $R \gtrsim 0.2 \REarth$.  Vertical lines indicate
-    the 90\% credible range.  We find $\Rpl = \rplrange$.}
-\end{figure}
-
-\begin{figure}
-  \includegraphics[width=\columnwidth]{bg}
-  \caption{\label{fig:det-bg} \textbf{Inferred detection probability
-      and density of background contamination.} (Top) The inferred
-    detection probability versus signal-to-noise ratio (see
-    Eq.\ \eqref{eq:pdetect}) from our parameterised model of selection
-    effects.  The solid line is the posterior median detection
-    probability and the shading gives the 90\% credible posterior
-    interval.  Our inferred detection probability is in rough
-    agreement with the measurements of detection efficiency in
-    Refs.\ \cite{Borucki2011,Batalha2013}.  (Bottom) The number
-    density of false-positive candidate signals, $d\Nbg/d\ln P \, d\ln
-    R$ (see Eq.\ \eqref{eq:background-rate}).  The density is highest
-    at large candidate period and radius, consistent with the dominant
-    source of contamination being background eclipsing
-    binaries\cite{Fressin2013}.  Overall, our model finds $\fposrange$
-    of the candidates are false-positive background signals,
-    consistent with the analysis in Ref.\ \cite{Fressin2013}.}
-\end{figure}
-
-\begin{figure}
-  \includegraphics[width=\columnwidth]{foreground-dist}
-  \caption{\label{fig:foreground-dist} \textbf{The inferred planet
-      period--radius distribution accounting for selection effects and
-      false-positives.}  (Upper Left) The planet number density per
-    logarithmic planet radius.  The density peaks at $\Rpeak =
-    \rpeakrange \REarth$ (90\% CL).  (Upper Right) The planet number
-    density in the period--radius plane.  The inferred correlation
-    coefficient between $\ln P$ and $\ln R$ is $r = \corrcoeffrange$.
-    (Lower Left) Scatter plot of the radius and period of the Kepler
-    planet candidates.  Color indicates the posterior false-positive
-    probability for each candidate.  Overall, the model prefers a
-    false-positive rate of $\fposrange$ (90\% CL).  The primary
-    contaminant is probably background eclipsing binaries; our
-    contamination rate is consistent with previous
-    work\cite{Fressin2013}. (Lower Right) The planet number density
-    per logarithmic planet period.  The density peaks at $P =
-    \ppeakrange \mathrm{yr}$ (90\% CL). }
-\end{figure}
-
-\begin{figure}
-  \includegraphics[width=\columnwidth]{selection}
-  \caption{\label{fig:selection} \textbf{Comparison of synthetic data
-      sets produced from the forward model incorporating selection
-      effects with observed candidates.}  (Upper Left) The observed
-    (black curve) and synthetic (blue curve; including planets and
-    false positives, and using the fitted selection model to
-    down-select the candidates from the planet distribution)
-    normalised candidate density per logarithmic radius.  Except for a
-    discrepancy at $R \simeq 10 \REarth$---associated with hot
-    Jupiters, a distinct planetary
-    population\cite{Albrecht2012,Naoz2012}---the model produces a good
-    fit to the observed candidates over the range of reported radii.
-    Note particularly the tail at large radii that comes from
-    background contaminants in both observed and synthetic data.
-    (Upper Right) Scatter plot of the observed candidates (black
-    circles) and a posterior-averaged draw of observed candidates from
-    the model (blue circles).  (Lower Left) Scatter plot of the
-    observed candidates.  Colors indicate the posterior-averaged
-    selection probability for each planet about its host star.  
-    The selection probability is treated a product of a geometric
-    factor giving the probability of an isotropically-oriented orbit
-    producing a transit and a signal-to-noise-ratio-dependent transit
-    detection probability.  (Lower Right) The observed (black curve)
-    and synthetic (blue curve; including planets and false positives,
-    and using the fitted selection model to down-select the candidates
-    from the planet distribution) normalised candidate density per
-    logarithmic period.  Except for the aforementioned hot Jupiter
-    peak at $P \simeq 1 \mathrm{day}$ the model produces a good fit to
-    the observed candidates over the range of reported periods.}
-\end{figure}
-
 \bibliography{kepler-selection}
 
-\begin{addendum}
+\begin{scilastnote}
 \item The code implementing this analysis is available under an
@@
   NASA Science Mission directorate.
-\item [Author Contributions] All authors assisted in the computational
-  modelling, discussed the results, and edited the manuscript.
-\item [Reprints] Reprints and permissions information is available at
-  \url{www.nature.com/reprints}.
-\item[Competing Interests] The authors declare that they have no
-  competing financial interests.
-\item[Correspondence] Correspondence and requests for materials should
-  be addressed to W.M.F.\ (email: \email{w.farr@bham.ac.uk}).
-\end{addendum}
+\end{scilastnote}
 

### 2014-11-26 code/plots.py

@@ -8,18 +8,18 @@ import scipy.stats as ss
 import triangle as tri
 
 def setup():
-    fig_width_pt = 469.75502
+    fig_width_pt = 455.24408
     inches_per_pt = 1.0/72.27
 
     fig_size = [fig_width_pt*inches_per_pt,
                 fig_width_pt*inches_per_pt]
 
     params = { 'backend' : 'pdf',
-               'axes.labelsize' : 10,
-               'font.size' : 10,
-               'legend.fontsize' : 10,
-               'xtick.labelsize' : 8,
-               'ytick.labelsize' : 8,
+               'axes.labelsize' : 12,
+               'font.size' : 12,
+               'legend.fontsize' : 12,
+               'xtick.labelsize' : 10,
+               'ytick.labelsize' : 10,
                'text.usetex' : True,
                'figure.figsize' : fig_size,
                'figure.autolayout' : True,
@@ -463,6 +463,9 @@ def plot_parameters(logpost, chain, eta_earths, outdir=None):
 def paper_plots(logpost, chain, eta_earths, outdir):
     pp.figure()
     plot_parameters(logpost, chain, eta_earths, outdir=outdir)
+
+    pp.figure()
+    plot_selection_background(logpost, chain, outdir=outdir)
 
     pp.figure()
     plot_foreground_distributions(logpost, chain, outdir=outdir)

### 2014-11-28 paper/cover-letter.tex
Some bugfixes, as submitted to Science.

@@
 
-\begin{letter}{Dr. Leslie Sage\\The Macmillan Building\\4 Crinan Street\\London N1 9XW\\United Kingdom}
+\begin{letter}{Dr. Maria Cruz\\Science/AAAS\\1200 New York Avenue NW\\Washington DC 20005\\United States}
 
-  \opening{Dear Dr. Sage:}
+  \opening{Dear Dr. Cruz:}
 
-  Please find enclosed a submission for Nature Letters entitled ``The
-  Occurrence of Earth-Like Planets Around Other Stars.''  The
-  submission comprises approximately 1900 words of text and four
-  figures (with captions).  We do not at this time have any extended
-  sections, supplementary information or supporting manuscripts.  The
-  requested brief paragraphs justifying publication in Nature and
-  summarising the letter's appeal to a broad audience follow.
+  Please find enclosed a report for submission to Science entitled
+  ``The Occurrence of Earth-Like Planets Around Other Stars.''  The
+  submission comprises approximately 2500 words of text, including
+  four figures (with captions).
 
-  \textbf{For the editor:} The quantity $\eta_\oplus$ is of broad
-  scientific interest.  Its estimation is a chief goal of NASA's
-  Kepler mission, but is complicated by the difficulty of detecting
-  Earth-like planets about other stars.  In addition to constraining
-  $\eta_\oplus$ using the complete Kepler data set, we have modelled
-  the distribution of planet orbital periods and radii and find that
-  it is well-fit by a single Gaussian component.  Our approach to
-  reconstructing the period-radius distribution of planets is
-  generally applicable to any survey with significant selection
-  effects.
-
-  \textbf{For the public:} We use the latest results from NASA's
-  Kepler mission to count how many planets of roughly Earth's size
-  orbit stars like the Sun in our galaxy at about the same distance
-  from their host stars as Earth.  We find that about 4\% of stars
-  have a planet the size of Earth in an Earth-like orbit; with about
-  100 billion Sun-like stars in the galaxy, this means there are over
-  4 billion Earth-sized planets in one year orbits in our galaxy!
+  The main result of our paper is a measurement of $\eta_\oplus$, the
+  number density of Earth-sized planets in Earth-like orbits about
+  stars similar to the sun.  Our measurement uses the complete data
+  release from the final quarter of operation of the Kepler sattelite
+  (Q17).  The quantity $\eta_\oplus$ is of broad scientific interest.
+  Its estimation is a chief goal of NASA's Kepler mission, but is
+  complicated by the difficulty of detecting Earth-like planets about
+  other stars (i.e.\ selection effects); we deal with this difficulty
+  using an innovative statistical technique that simultaneously fits
+  the intrinsic distribution of planet periods and radii and a
+  parameterised model of the Kepler selection function.  In addition
+  to constraining $\eta_\oplus$, we have modelled the distribution of
+  planet orbital periods and radii and find that it is well-fit by a
+  single Gaussian component.  Our approach to reconstructing the
+  intrinsic distribution of planet periods and radii is generally
+  applicable to any survey with significant selection effects.
 

### 2014-11-28 paper/kepler-selection.tex
Some bugfixes, as submitted to Science.

@@
 orbits.  As we shall see (Figure\ \ref{fig:selection}), this simple
-model combined with the aforementioned selection function, provides a
+model combined with the aforementioned selection function provides a
 good fit to the Kepler candidate distribution.  In our model, observed
@@
     \Sigma_{RR} \right]\right]$ the covariance matrix of $\ln P$ and
-$\ln R$; $N\left[ \mu, \Sigma \right](x,y)$ is the normal distribution
-Our model assumes that planets appear around their host stars in a
-Poisson process; this is almost certainly wrong in
+$\ln R$; $N\left[ \mu, \Sigma \right](x,y)$ is the normal
+distribution.  Our model assumes that planets appear around their host
+stars in a Poisson process; this is almost certainly wrong in
 detail\cite{Weissbein2012}, but nevertheless provides a good fit to
@@
   \caption{\label{fig:eta-earth} \textbf{Posteriors on $\etaearth$ and
-      $Rpl$ accounting for selection effects and false-positive
+      $\Rpl$ accounting for selection effects and false-positive
       detections.}  (Top) The inferred posterior density on $\etaearth

### 2014-11-28 code/plots.py

@@ -1,9 +1,13 @@
+#!/usr/bin/env python
+import bz2
+import data_processing as dp
 import matplotlib.colors as mc
 import matplotlib.pyplot as pp
 import numpy as np
 import os.path as op
 import plotutils.autocorr as ac
 import plotutils.plotutils as pu
+import posterior as pos
 import scipy.stats as ss
 import triangle as tri
 
@@ -323,8 +327,8 @@ def plot_nplanets_histogram(chain, outdir=None):
         nps = chain[:,::tau,0].flatten()
 
     pu.plot_histogram_posterior(nps, normed=True, color='k', histtype='step')
-    pp.xlabel(r'$R_\mathrm{pl}$')
-    pp.ylabel(r'$p(R_\mathrm{pl})$')
+    pp.xlabel(r'$\Lambda_\mathrm{pl}$')
+    pp.ylabel(r'$p(\Lambda_\mathrm{pl})$')
 
     pp.axvline(np.percentile(nps, 5), color='k')
     pp.axvline(np.percentile(nps, 95), color='k')
@@ -451,8 +455,8 @@ def plot_parameters(logpost, chain, eta_earths, outdir=None):
     pp.subplot(2,1,2)
     nps = chain[:,:,0].flatten()
     pu.plot_histogram_posterior(nps, normed=True, color='k', histtype='step')
-    pp.xlabel(r'$R_\mathrm{pl}$')
-    pp.ylabel(r'$p(R_\mathrm{pl})$')
+    pp.xlabel(r'$\Lambda_\mathrm{pl}$')
+    pp.ylabel(r'$p(\Lambda_\mathrm{pl})$')
 
     pp.axvline(np.percentile(nps, 5), color='k')
     pp.axvline(np.percentile(nps, 95), color='k')
@@ -494,3 +498,17 @@ def paper_tex(logpost, chain, eta_earths, rs, outdir):
 
         out.write('\\newcommand{{\\ppeakrange}}{{{0:0.3f}_{{-{1:0.3f}}}^{{+{2:0.3f}}}}}\n'.format(pp[0], pp[1], pp[2]))
         
+def do_it_all(outdir):
+    with bz2.BZ2File('../../Kepler/chain.npy.bz2', 'r') as inp:
+        chain = np.load(inp)
+    tchain = ac.emcee_thinned_chain(chain)
+    candidates = dp.load_candidate_properties('../../Kepler/candidates.dat.bz2')
[…]

### 2014-12-15 paper/kepler-selection.tex
Re-tooled for ApJL (sigh).

@@
-\documentclass[12pt]{article}
-
-\usepackage{scicite}
-\usepackage{times}
-
-\topmargin 0.0cm
-\oddsidemargin 0.2cm
-\textwidth 16cm 
-\textheight 21cm
-\footskip 1.0cm
-
-\newenvironment{sciabstract}{%
-\begin{quote} \bf}
-{\end{quote}}
-
-\renewcommand\refname{References and Notes}
-
-\newcounter{lastnote}
-\newenvironment{scilastnote}{%
-\setcounter{lastnote}{\value{enumiv}}%
-\addtocounter{lastnote}{+1}%
-\begin{list}%
-{\arabic{lastnote}.}
-{\setlength{\leftmargin}{.22in}}
-{\setlength{\labelsep}{.5em}}}
-{\end{list}}
+\documentclass[apjl]{emulateapj}
 
@@
 \usepackage{amssymb}
-\usepackage{xcolor}
 \usepackage{graphicx}
-
-\bibliographystyle{Science}
 
@@
 
-\newcommand{\email}[1]{\href{mailto:#1}{\nolinkurl{#1}}}
-
 \newcommand{\Will}[1]{\textcolor{cyan}{Will: #1}}
@@
 
-\newcommand{\aj}{The Astronomical Journal}
-\newcommand{\aap}{Astronomy \& Astrophysics}
-\newcommand{\apj}{The Astrophysical Journal}
-\newcommand{\apjl}{The Astrophysical Journal Letters}
-\newcommand{\apjs}{The Astrophysical Journal Supplement}
-\newcommand{\mnras}{Monthly Notices of the Royal Astronomical Society}
-\newcommand{\pasp}{Publications of the Astronomical Society of the Pacific}
+%% \newcommand{\aj}{The Astronomical Journal}
+%% \newcommand{\aap}{Astronomy \& Astrophysics}
+%% \newcommand{\apj}{The Astrophysical Journal}
+%% \newcommand{\apjl}{The Astrophysical Journal Letters}
+%% \newcommand{\apjs}{The Astrophysical Journal Supplement}
+%% \newcommand{\mnras}{Monthly Notices of the Royal Astronomical Society}
+%% \newcommand{\pasp}{Publications of the Astronomical Society of the Pacific}
 
@@
 
-\title{The Occurrence of Earth-Like Planets Around Other Stars}
-\author{Will M. Farr$^{1\ast}$, Ilya Mandel$^{1}$, Chris Aldridge$^{1}$ \& Kirsty Stroud$^{1}$\\
-  \normalsize{$^1$School of Physics and Astronomy}\\
-  \normalsize{University of Birmingham}\\
-  \normalsize{Birmingham}\\
-  \normalsize{B15 2TT}\\
-  \normalsize{United Kingdom}\\
-  \normalsize{$^\ast$To whom correspondence should be addressed; E-mail: w.farr@bham.ac.uk.}}
-
-\date{}
-
 \begin{document}
 
-\baselineskip24pt
+\title{The Occurrence of Earth-Like Planets Around Other Stars}
+\shorttitle{Earth-Like Planets Around Other Stars}
+\author{Will M. Farr, Ilya Mandel, Chris Aldridge \and Kirsty Stroud}
+\affil{School of Physics and Astronomy\\University of
+  Birmingham\\Birmingham\\B15 2TT\\United Kingdom}
+\email{w.farr@bham.ac.uk}
+\email{ilyamandel@chgk.info}
+\email{cxa064@bham.ac.uk}
+\email{kls081@bham.ac.uk}
+
+\begin{abstract}
+  The quantity $\etaearth$, the number density of planets per star per
+  logarithmic planetary radius per logarithmic orbital period at one
+  Earth radius and one year period, describes the occurrence of
+  Earth-like extrasolar planets.  Here we present a measurement of
+  $\etaearth$ from a parameterised forward model of the (correlated)
+  period-radius distribution and the observational selection function
+  in the most recent (Q17) data release from the Kepler satellite.  We
+  find $\etaearth = \earange$ (90\% CL).  We conclude that each star
+  hosts $\rplrange$ planets with $P \lesssim 3 \mathrm{yr}$ and $R
+  \gtrsim 0.2 \REarth$.  Our empirical model for false-positive
+  contamination is consistent with the dominant source being
+  background eclipsing binary stars.  The distribution of planets we
+  infer is consistent with a highly-stochastic planet formation
+  process producing many correlated, fractional changes in planet
+  sizes and orbits.
+\end{abstract}
+
+\keywords{planetary systems---planets and satellites: fundamental
+  parameters---planets and satellites: detection---methods:
+  statistical}
 
@@
 
-\begin{sciabstract}
-  The quantity $\etaearth$, the number density of planets per star per
-  logarithmic planetary radius per logarithmic orbital period at one
-  Earth radius and one year period, describes the occurrence of
-  Earth-like extrasolar planets.  Here we present a
-  measurement of $\etaearth$ from a parameterised forward model of the
-  (correlated) period-radius distribution and the observational
-  selection function in the most recent (Q17) data release from the
-  Kepler satellite\cite{Borucki2010,Borucki2011,Batalha2013}.  We find
-  $\etaearth = \earange$ (90\% CL).  We conclude that each star hosts
-  $\rplrange$ planets with $P \lesssim 3 \mathrm{yr}$ and $R \gtrsim
-  0.2 \REarth$.  Our empirical model for false-positive contamination
-  is consistent with the dominant source being background eclipsing
-  binary stars\cite{Fressin2013}.  The distribution of planets we
-  infer is consistent with a highly-stochastic planet formation
-  process producing many correlated, fractional changes in planet
-  sizes and orbits.  
-\end{sciabstract}
+\section{Introduction}
 
@@
 period-radius distribution and the observational selection function in
-the most recent (Q17) data release from the Kepler
-satellite\cite{Borucki2010,Borucki2011,Batalha2013}.  Our data set
-comprises 181,568 systems observed under the Kepler exoplanet
-observing program (mostly G-type stars on the main
-sequence\cite{Batalha2010}), producing 2598 planetary candidates.  We
-parameterise the distribution of planetary periods and radii using a
-single, correlated Gaussian component; treat selection effects using a
-parameterised transit detection probability based on the measured
-noise level and stellar properties in the Kepler catalog; and include
-an empirically-parameterised, independent component in the
-period-radius distribution to represent false-positive planet
-detections.  Using our model we can simultaneously estimate
-$\etaearth$, place constraints on the planet period-radius
-distribution function, and determine the degree of contamination by
-false-positive candidate identifications.  We find $\etaearth =
-\earange$ (90\% CL).  We conclude that each star hosts $\rplrange$
-planets with $P \lesssim 3 \mathrm{yr}$ and $R \gtrsim 0.2 \REarth$,
-that the peak of the planet radius distribution lies at $\Rpeak =
-\rpeakrange \REarth$, and that $\ln P$ and $\ln R$ are correlated with
-correlation coefficient $r = \corrcoeffrange$ (all 90\% CL).  Our
-empirical model for false-positive contamination is consistent with
-the dominant source being background eclipsing binary
-stars\cite{Fressin2013}, with $\fposrange$ (90\% CL) of the candidates
-being false-positives.  The distribution of planets we infer is
-consistent with a highly-stochastic planet formation process producing
-many correlated, fractional changes in planet sizes and orbits.  Our
-approach of determining both the intrinsic distribution of objects and
-selection effects empirically from survey data is generally
-applicable.
+the most recent (Q17) data release from the Kepler satellite
+\citep{Borucki2010,Borucki2011,Batalha2013}.  Our data set comprises
+181,568 systems observed under the Kepler exoplanet observing program
+(mostly G-type stars on the main sequence \citep{Batalha2010}),
+producing 2598 planetary candidates.  We parameterise the distribution
+of planetary periods and radii using a single, correlated Gaussian
+component; treat selection effects using a parameterised transit
+detection probability based on the measured noise level and stellar
+properties in the Kepler catalog; and include an
+empirically-parameterised, independent component in the period-radius
+distribution to represent false-positive planet detections.  Using our
+model we can simultaneously estimate $\etaearth$, place constraints on
+the planet period-radius distribution function, and determine the
+degree of contamination by false-positive candidate identifications.
+We find $\etaearth = \earange$ (90\% CL).  We conclude that each star
+hosts $\rplrange$ planets with $P \lesssim 3 \mathrm{yr}$ and $R
+\gtrsim 0.2 \REarth$, that the peak of the planet radius distribution
+lies at $\Rpeak = \rpeakrange \REarth$, and that $\ln P$ and $\ln R$
+are correlated with correlation coefficient $r = \corrcoeffrange$ (all
+90\% CL).  Our empirical model for false-positive contamination is
+consistent with the dominant source being background eclipsing binary
+stars \citep{Fressin2013}, with $\fposrange$ (90\% CL) of the
+candidates being false-positives.  The distribution of planets we
+infer is consistent with a highly-stochastic planet formation process
+producing many correlated, fractional changes in planet sizes and
+orbits.  Our approach of determining both the intrinsic distribution
+of objects and selection effects empirically from survey data is
+generally applicable.
 
@@
 team from observations of stars in the ``EX'' observing program (which
-are primarily G-type main-sequence stars similar to our own
-Sun\cite{Batalha2010}), giving the inferred planetary period and
-radius for each.  The fractional depth of a planetary transit signal
-depends only on the radii of the planet and its host star.  The signal
-to noise ratio of a series of transits about a particular star in the
+are primarily G-type main-sequence stars similar to our own Sun
+\citep{Batalha2010}), giving the inferred planetary period and radius
+for each.  The fractional depth of a planetary transit signal depends
+only on the radii of the planet and its host star.  The signal to
+noise ratio of a series of transits about a particular star in the
 Kepler satellite scales with planetary period and radius
-as\cite{Chatterjee2012}
+as \citep{Chatterjee2012}
 \begin{equation}
@@
 quarters of observation of that star, the stellar radius and mass, and
-the intrinsic variability of the stellar
-intensity\cite{Christiansen2012}.  In our analysis, we obtain these
-quantities from the Kepler Input Catalog\cite{Batalha2010,Brown2011}
-and the MAST Kepler
-archive\footnote{\url{http://archive.stsci.edu/kepler/}}.  To a good
-approximation (see Fig.\ \ref{fig:selection} below), the detectability
-of a series of planetary transits in the Kepler data set is a function
-of the signal to noise ratio of the series.  Because the detectability
-of planet transits depends on both period and radius, it is important
-to consider the joint (i.e., two-dimensional) distribution of these
-quantities in the data\cite{Tabachnik2002,Youdin2011}.  We model the
-detection probability of a transit as a function that rises linearly
-in the log of the signal to noise ratio from zero at a threshold
-signal to noise to one at a larger signal to noise:
+the intrinsic variability of the stellar intensity
+\citep{Christiansen2012}.  In our analysis, we obtain these quantities
+from the Kepler Input Catalog \citep{Batalha2010,Brown2011} and the
+MAST Kepler archive\footnote{\url{http://archive.stsci.edu/kepler/}}.
+
+\section{Model}
+
+To a good approximation (see Fig.\ \ref{fig:selection} below), the
+detectability of a series of planetary transits in the Kepler data set
+is a function of the signal to noise ratio of the series.  Because the
+detectability of planet transits depends on both period and radius, it
+is important to consider the joint (i.e., two-dimensional)
+distribution of these quantities in the data
+\citep{Tabachnik2002,Youdin2011}.  We model the detection probability
+of a transit as a function that rises linearly in the log of the
+signal to noise ratio from zero at a threshold signal to noise to one
+at a larger signal to noise:
 \begin{equation}
@@
 $\rhomin = \rhominrange$ and $\rhomax = \rhomaxrange$ (90\% CL), in
-rough agreement with Refs.\ \cite{Borucki2011,Batalha2013}.  A plot of
-our inferred detection probability appears in Fig.\ \ref{fig:det-bg}
+rough agreement with \citet{Borucki2011,Batalha2013}.  A plot of our
+inferred detection probability appears in Fig.\ \ref{fig:det-bg}
 
@@
     agreement with the measurements of detection efficiency in
-    Refs.\ \cite{Borucki2011,Batalha2013}.  (Bottom) The number
-    density of false-positive candidate signals, $d\Nbg/d\ln P \, d\ln
-    R$ (see Eq.\ \eqref{eq:background-rate}).  The density is highest
-    at large candidate period and radius, consistent with the dominant
+    \citet{Borucki2011,Batalha2013}.  (Bottom) The number density of
+    false-positive candidate signals, $d\Nbg/d\ln P \, d\ln R$ (see
+    Eq.\ \eqref{eq:background-rate}).  The density is highest at small
+    candidate period and large radius, consistent with the dominant
     source of contamination being background eclipsing
-    binaries\cite{Fressin2013}.  Overall, our model finds $\fposrange$
-    of the candidates are false-positive background signals,
-    consistent with the analysis in Ref.\ \cite{Fressin2013}.}
+    binaries \citep{Fressin2013}.  Overall, our model finds
+    $\fposrange$ of the candidates are false-positive background
+    signals, consistent with the analysis in \citet{Fressin2013}.}
 \end{figure}
 
-The probability that a planet's orbital plane will align with the line-of-sight to
-Earth and thereby produce a transit signal is
+The probability that a planet's orbital plane will align with the
+line-of-sight to Earth and thereby produce a transit signal is
 \begin{equation}
@@
 planets populate the candidate $P$-$R$ plane with number density
-\begin{equation}
+\begin{multline}
   \label{eq:foreground-rate}
   \frac{dN_\mathrm{obs}}{d\ln P\, d\ln R} = \left[ \sum_\mathrm{stars}
-    p_\mathrm{select}(P, R) \right] \Rpl N\left[ \mu, \Sigma
+    p_\mathrm{select}(P, R) \right] \\ \times \Rpl N\left[ \mu, \Sigma
     \right]\left( \ln P, \ln R \right),
-\end{equation}
+\end{multline}
 where $\Rpl$, $\mu$, and $\Sigma$ are parameters of our model, with
@@
 distribution.  Our model assumes that planets appear around their host
-stars in a Poisson process; this is almost certainly wrong in
-detail\cite{Weissbein2012}, but nevertheless provides a good fit to
-the observed data (see Figure \ref{fig:selection}).
+stars in a Poisson process; this is almost certainly wrong in detail
+\citep{Weissbein2012}, but nevertheless provides a good fit to the
+observed data (see Figure \ref{fig:selection}).
 
@@
 gradient across a rectangular region in the $\ln P$-$\ln R$ plane:
-\begin{equation}
+\begin{multline}
   \label{eq:background-rate}
   \frac{d\Nbg}{d \ln P \, d \ln R} = \frac{\Nbg}{\Delta \ln P \Delta
-    \ln R} \left( 1 + \vec{\gamma} \cdot \left[ \ln P - \ln P_\mathrm{mid} , \ln
+    \ln R} \\ \times \left( 1 + \vec{\gamma} \cdot \left[ \ln P - \ln P_\mathrm{mid} , \ln
     R - \ln R_\mathrm{mid} \right] \right),
-\end{equation}
+\end{multline}
 where $\Delta \ln P = \ln P_\mathrm{max} - \ln P_\mathrm{min}$, $\ln
@@
 purely empirical model for the background contamination, but is
-reasonable if the chief contaminant is background eclipsing
-binaries\cite{Fressin2013,Duquennoy1991}.  The posterior on the
-background number density in the $P$-$R$ plane appears in
+reasonable if the chief contaminant is background eclipsing binaries
+\citep{Fressin2013,Duquennoy1991}.  The posterior on the background
+number density in the $P$-$R$ plane appears in
 Figure\ \ref{fig:det-bg}.
 
-Unlike Ref.\ \cite{Foreman-Mackey2014}, we do not attempt to model the
+Unlike \citet{Foreman-Mackey2014}, we do not attempt to model the
 observational uncertainties in the estimated periods and radii from
 the Kepler candidate data set.  In spite of several candidates with
-very large uncertainties in measured parameters, we have found that our
-fit is essentially unchanged when applied to synthetic observations
-with periods and radii re-drawn from the range of observational
-uncertainties quoted in the Q17 data release.
+very large uncertainties in measured parameters, we have found that
+our fit is essentially unchanged when applied to synthetic
+observations with periods and radii re-drawn from the range of
+observational uncertainties quoted in the Q17 data release.
 
 The likelihood of the observed periods and radii under our model is an
-inhomogeneous Poisson likelihood\cite{Farr2013,Youdin2011} with a rate
-that is the sum of Eq.\ \eqref{eq:foreground-rate} and
+inhomogeneous Poisson likelihood \citep{Farr2013,Youdin2011} with a
+rate that is the sum of Eq.\ \eqref{eq:foreground-rate} and
 Eq.\ \eqref{eq:background-rate}.  We impose priors on our 15 model
@@
 (implicitly) the parameters describing selection effects, we impose a
-$\frac{1}{\sqrt{\Npl}}$ prior; for the background rate $\Rbg$ we
-impose a $\frac{1}{\sqrt{\Rbg}}$ prior; for the selection model
-parameters $\rhomin$ and $\rhomax$ we impose a log-normal prior with
-unit width at signal to noise ratios of 3 and 11, respectively; in all
-other parameters we impose a flat (i.e., constant-density) prior.  The
-product of likelihood and prior gives a Bayesian posterior density
-function on the fifteen-dimensional parameter space of our model.  We
-sample from this function using the \texttt{emcee}
-sampler\cite{Foreman-Mackey2013}.  The posterior describes
-simultaneously the intrinsic distribution and number of exoplanets,
-the amount and distribution of the contaminating false-positive events
-in the candidate data set, and the selection function of the
-instrument for true planetary transit events.
+$1/\sqrt{\Npl}$ prior; for the background rate $\Rbg$ we impose a
+$1/\sqrt{\Rbg}$ prior; for the selection model parameters $\rhomin$
+and $\rhomax$ we impose a log-normal prior with unit width at signal
+to noise ratios of 3 and 11, respectively; in all other parameters we
+impose a flat (i.e., constant-density) prior.  The product of
+likelihood and prior gives a Bayesian posterior density function on
+the fifteen-dimensional parameter space of our model.  We sample from
+this function using the \texttt{emcee} sampler
+\citep{Foreman-Mackey2013}.  The posterior describes simultaneously
+the intrinsic distribution and number of exoplanets, the amount and
+distribution of the contaminating false-positive events in the
+candidate data set, and the selection function of the instrument for
+true planetary transit events.
+
+\section{Conclusion}
 
@@
 Recall that
-\begin{equation}
+\begin{multline}
   \etaearth = \left. \frac{dN}{d \ln P \ln R} \right|_{R = \REarth, P
-    = 1\,\mathrm{yr}} = \Rpl N\left[ \mu, \Sigma \right]\left( \ln 1\,\mathrm{yr},
-  \ln \REarth \right),
-\end{equation}
+    = 1\,\mathrm{yr}} \\ = \Rpl N\left[ \mu, \Sigma \right]\left( \ln
+  1\,\mathrm{yr}, \ln \REarth \right),
+\end{multline}
 which is roughly the number of planets per star with periods and radii
@@
 finds that the false-positive rate in the candidate data set is
-$\fposrange$ (90\% CL), consistent with previous
-work\cite{Fressin2013} estimating the contamination in the Kepler
+$\fposrange$ (90\% CL), consistent with previous work
+\citep{Fressin2013} estimating the contamination in the Kepler
 candidate set.  Our model has the peak of the planet period-radius
@@
     contaminant is probably background eclipsing binaries; our
-    contamination rate is consistent with previous
-    work\cite{Fressin2013}. (Lower Right) The planet number density
-    per logarithmic planet period.  The density peaks at $P =
-    \ppeakrange \mathrm{yr}$ (90\% CL). }
+    contamination rate is consistent with previous work
+    \citep{Fressin2013}. (Lower Right) The planet number density per
+    logarithmic planet period.  The density peaks at $P = \ppeakrange
+    \mathrm{yr}$ (90\% CL). }
 \end{figure}
@@
 with the already-observed candidate set.  These predictions can be
-used to perform graphical and posterior-predictive model
-checking\cite{Gelman2013}.  Fig.\ \ref{fig:selection} compares the
+used to perform graphical and posterior-predictive model checking
+\citep{Gelman2013}.  Fig.\ \ref{fig:selection} compares the
 predictions of our model for observed periods and radii (incorporating
@@
 the selection function of the Kepler telescope and pipeline.  Except
-for the known sub-population of hot
-Jupiters\cite{Albrecht2012,Naoz2012}, our model provides a very good
-fit to the observed data.  That a simple log-normal distribution in
-period and radius fits the observed distribution of planets well may
-indicate that planet formation is a stochastic process with many
-small, correlated, and multiplicative influences on planet period and
-radius resulting, from the central limit theorem, in a log-normal
+for the known sub-population of hot Jupiters
+\citep{Albrecht2012,Naoz2012}, our model provides a very good fit to
+the observed data.  That a simple log-normal distribution in period
+and radius fits the observed distribution of planets well may indicate
+that planet formation is a stochastic process with many small,
+correlated, and multiplicative influences on planet period and radius
+resulting, from the central limit theorem, in a log-normal
 distribution in these parameters.
@@
     \simeq 10 \REarth$---associated with hot Jupiters, a distinct
-    planetary population\cite{Albrecht2012,Naoz2012}---the model
+    planetary population \citep{Albrecht2012,Naoz2012}---the model
     produces a good fit to the observed candidates over the range of
@@
 
-Previous
-estimates\cite{Catanzarite2011,Traub2012,Dong2013,Petigura2013,Foreman-Mackey2014}
+Previous estimates
+\citep{Catanzarite2011,Traub2012,Dong2013,Petigura2013,Foreman-Mackey2014}
 place $1\% \lesssim \etaearth \lesssim 34\%$.  These works dealt with
@@
 region of the period-radius parameter space where observations are
-complete and extrapolating to $R = \REarth$ and $P = 1
-\mathrm{yr}$\cite{Catanzarite2011,Traub2012}, applying a binned
-analysis incorporating survey incompleteness in the period-radius
-plane\cite{Dong2013,Petigura2013} or analysing the results of a
-customised planet detection pipeline on a subset of the Kepler
-observations\cite{Petigura2013,Foreman-Mackey2014}.  The methods and
-analysed data sets of Refs.\ \cite{Petigura2013,Foreman-Mackey2014}
-are most comparable to ours.  These studies used the same data set,
-produced\cite{Petigura2013} from a subset of the available Kepler data
-and a customised pipeline to search for transit signals.  They both
+complete and extrapolating to $R = \REarth$ and $P = 1 \mathrm{yr}$
+\citep{Catanzarite2011,Traub2012}, applying a binned analysis
+incorporating survey incompleteness in the period-radius plane
+\citep{Dong2013,Petigura2013} or analysing the results of a customised
+planet detection pipeline on a subset of the Kepler observations
+\citep{Petigura2013,Foreman-Mackey2014}.  The methods and analysed
+data sets of \citet{Petigura2013,Foreman-Mackey2014} are most
+comparable to ours.  These studies used the same data set, produced
+\citep{Petigura2013} from a subset of the available Kepler data and a
+customised pipeline to search for transit signals.  They both
 accounted for selection effects by measuring the recoverability of
@@
 Our result for $\etaearth$ is consistent with, but more precise than,
-Ref.\ \cite{Foreman-Mackey2014} and (somewhat) inconsistent with
-Ref.\ \cite{Petigura2013}.
+\citet{Foreman-Mackey2014} and (somewhat) inconsistent with
+\citet{Petigura2013}.
+
+\acknowledgements
+
+The code implementing this analysis is available under an open-source
+``MIT'' license at \url{https://github.com/farr/kepler-selection}.
+This work was supported by the Science and Technology Facilities
+Council.  Computations in this work were performed on the University
+of Birmingham's BlueBEAR cluster.  Some of the data presented in this
+paper were obtained from the Mikulski Archive for Space Telescopes
+(MAST). STScI is operated by the Association of Universities for
+Research in Astronomy, Inc., under NASA contract NAS5-26555. Support
+for MAST for non-HST data is provided by the NASA Office of Space
+Science via grant NNX13AC07G and by other grants and contracts.  This
+paper includes data collected by the Kepler mission. Funding for the
+Kepler mission is provided by the NASA Science Mission directorate.
 
@@
 
-\begin{scilastnote}
-\item The code implementing this analysis is available under an
-  open-source ``MIT'' license at
-  \url{https://github.com/farr/kepler-selection}.  This work was
-  supported by the Science and Technology Facilities Council.
-  Computations in this work were performed on the University of
-  Birmingham's BlueBEAR cluster.  Some of the data presented in
-  this paper were obtained from the Mikulski Archive for Space
-  Telescopes (MAST). STScI is operated by the Association of
-  Universities for Research in Astronomy, Inc., under NASA contract
-  NAS5-26555. Support for MAST for non-HST data is provided by the
-  NASA Office of Space Science via grant NNX13AC07G and by other
-  grants and contracts.  This paper includes data collected by the
-  Kepler mission. Funding for the Kepler mission is provided by the
-  NASA Science Mission directorate.
-\end{scilastnote}
-
 \end{document}

### 2014-12-15 code/plots.py

@@ -12,18 +12,18 @@ import scipy.stats as ss
 import triangle as tri
 
 def setup():
-    fig_width_pt = 455.24408
+    fig_width_pt = 245.26653
     inches_per_pt = 1.0/72.27
 
     fig_size = [fig_width_pt*inches_per_pt,
                 fig_width_pt*inches_per_pt]
 
     params = { 'backend' : 'pdf',
-               'axes.labelsize' : 12,
-               'font.size' : 12,
-               'legend.fontsize' : 12,
-               'xtick.labelsize' : 10,
-               'ytick.labelsize' : 10,
+               'axes.labelsize' : 8,
+               'font.size' : 8,
+               'legend.fontsize' : 8,
+               'xtick.labelsize' : 6,
+               'ytick.labelsize' : 6,
                'text.usetex' : True,
                'figure.figsize' : fig_size,
                'figure.autolayout' : True,
@@ -242,6 +242,8 @@ def plot_foreground_distributions(logpost, chain, N=1000, outdir=None):
     pp.xlabel(r'$P$ ($\mathrm{yr}$)')
     pp.ylabel(r'$dN/d\ln P$')
 
+    pp.tight_layout()
+    
     if outdir is not None:
         pp.savefig(op.join(outdir, 'foreground-dist.pdf'))
         
@@ -315,6 +317,8 @@ def plot_selection(logpost, chain, Ndraw=100, outdir=None):
     pp.xlabel(r'$P$ ($\mathrm{yr}$)')
     pp.ylabel(r'$p(\ln P)$')
 
+    pp.tight_layout()
+    
     if outdir is not None:
         pp.savefig(op.join(outdir, 'selection.pdf'))
     
@@ -438,6 +442,8 @@ def plot_selection_background(logpost, chain, outdir=None):
     pp.ylabel(r'$R$ ($R_\oplus$)')
 
     pp.axis(xmin=np.exp(ll[0]), ymin=np.exp(ll[1]), xmax=np.exp(ur[0]), ymax=np.exp(ur[1]))
+
+    pp.tight_layout()
     
     if outdir is not None:
         pp.savefig(op.join(outdir, 'bg.pdf'))
@@ -460,6 +466,8 @@ def plot_parameters(logpost, chain, eta_earths, outdir=None):
 
     pp.axvline(np.percentile(nps, 5), color='k')
     pp.axvline(np.percentile(nps, 95), color='k')
+
[…]

### 2014-12-16 code/plots.py
Added code to produce FITS files to go along with figures for DbF.

@@ -1,4 +1,5 @@
 #!/usr/bin/env python
+from astropy.io import fits
 import bz2
 import data_processing as dp
 import matplotlib.colors as mc
@@ -246,6 +247,27 @@ def plot_foreground_distributions(logpost, chain, N=1000, outdir=None):
     
     if outdir is not None:
         pp.savefig(op.join(outdir, 'foreground-dist.pdf'))
+
+        hdr = fits.Header()
+        hdr['COMMENT'] = 'Farr et al (2015) Figure 3'
+        hdr['COMMENT'] = 'Posterior mean on the true foreground number density'
+        hdr['COMMENT'] = 'and posterior false-alarm probability for all candidates'
+
+        hdu = fits.PrimaryHDU(header=hdr)
+        hdulist = fits.HDUList([hdu])
+
+        pers = fits.Column(name='P', format='E', unit='yr', array=PS.flatten())
+        rads = fits.Column(name='R', format='E', unit='REarth', array=RS.flatten())
+        dens = fits.Column(name='dNdlnPdlnR', format='E', array=dNdPdR.flatten())
+        hdulist.append(fits.BinTableHDU.from_columns(fits.ColDefs([pers, rads, dens])))
+
+        names = fits.Column(name='ID', format='D', array=logpost.candidates['Kepler_ID'])
+        pers = fits.Column(name='P', format='E', unit='yr', array=logpost.candidates['Period'])
+        rads = fits.Column(name='R', format='E', unit='REarth', array=logpost.candidates['Radius'])
+        faps = fits.Column(name='FAP', format='E', array=pbacks)
+        hdulist.append(fits.BinTableHDU.from_columns(fits.ColDefs([names, pers, rads, faps])))
+
+        hdulist.writeto(op.join(outdir, 'foreground-dist.fits'), clobber=True)
         
 def plot_selection(logpost, chain, Ndraw=100, outdir=None):
     fchain = chain.reshape((-1, chain.shape[2]))
@@ -321,7 +343,28 @@ def plot_selection(logpost, chain, Ndraw=100, outdir=None):
     
     if outdir is not None:
         pp.savefig(op.join(outdir, 'selection.pdf'))
-    
+
+        hdr = fits.Header()
+        hdr['COMMENT'] = 'Farr et al (2015) Figure 4'
+        hdr['COMMENT'] = 'Posterior predictive distribution of observed periods and radii'
[…]

### 2014-12-16 code/plots.py
Fixed figure mis-labeling in comments of FITS files (plots.py).

@@ -492,7 +492,7 @@ def plot_selection_background(logpost, chain, outdir=None):
         pp.savefig(op.join(outdir, 'bg.pdf'))
 
         hdr = fits.Header()
-        hdr['COMMENT'] = 'Farr et al (2015) Figure 2'
+        hdr['COMMENT'] = 'Farr et al (2015) Figure 1'
         hdr['COMMENT'] = 'Posterior on the selection function and contamination'
         hdr['COMMENT'] = 'log_rho_min and log_rho_max (natural logs) are parameters giving'
         hdr['COMMENT'] = 'lower and upper bounds of the linear-in-log part of s.f.'
@@ -557,7 +557,7 @@ def plot_parameters(logpost, chain, eta_earths, outdir=None):
         tbhdu = fits.BinTableHDU.from_columns(cols)
 
         hdr = fits.Header()
-        hdr['COMMENT'] = 'Farr, et al (2015) Figure 1'
+        hdr['COMMENT'] = 'Farr, et al (2015) Figure 2'
         hdr['COMMENT'] = 'Posterior samples from eta_earth and Lambda_pl'
         hdr['COMMENT'] = 'eta_earth is the number density of planets per star'
         hdr['COMMENT'] = 'per natural-logarithmic interval in P and R'

### 2015-03-10 paper/response.txt
Proposed changes for response to referee report.

For the editor
--------------

As you will see from our response to the referee attached, we have
identified a number of major errors in the referee's discussion of our
work. Additionally, the referee seems unusually hostile to our
modelling choices (e.g. "I could go on with many more effects that are
not adequately modeled, and often not even discussed. But it's not a
referee's job to develop a research plan for them.").  As you will see
in the response, we feel that our choices are well-justified.

Given these issues, could we request to have a second, independent
referee comment on our revised work?  We don't know the procedures for
this, but if appropriate we would be happy for both referees to see
this note to the editor and the attached referee response.

We have made some changes to the discussion in the paper in response
to the current referee's misunderstandings; those are outlined in the
course of the response.

Response
--------

There are several problems with the referee's main points of criticism
of our paper, and also several outright mistakes. Nevertheless, in
some cases we feel that the referee has valid criticisms of our paper.
Below we explain the changes we have made to our manuscript to attempt
to address the referee's criticism and where we feel that criticism is
in error.

The referee writes:

  The entire analysis is predicated on assuming that the intrinsic
  distribution for planets is described by a two-dimensional normal
  distribution and that a piece-wise linear approximation to the
  detection probability that depends only on the estimated transit
  signal to noise. The authors argue that this is a reasonable model,
  based primarily on high signal-to-noise detections of planets much
  larger than Earth and at orbital periods much less than one year.
  However, they they proceed to extrapolate to regions of planet
  radius-orbital period space for which there are very few viable planet
  candidates. Both of their headline results are based on such an
  extrapolation.

The referee is incorrect that our model validation is based on high
signal-to-noise planets at large radii and short periods.  As is
clearly shown in Figure 4 of our manuscript, synthetic data drawn from
our fitted model matches the observations well at:

 * Small planet radius irrespective of period (upper left panel)
 * Peak of the fitted radius distribution (R = 1.25 REarth) (upper
   left panel)
 * Large radius, where the candidates are dominated by false-positives
   (upper left panel)
 * Small periods (lower right panel)
 * Peak of the period distribution (lower right panel)
 * Large period (lower right panel)

It is simply incorrect to say that our model is only validated by
comparison to candidates at much larger than Earth radius and shorter
than one-year periods. The referee *is* correct that there are
relatively few candidates in the P-R plane near 1 year and 1 REarth,
but this is exactly why it is important to validate the model in a
wide range of periods and radii that span the 1 year - 1 REarth region
of interest. We admit to being surprised that such a simple model
reproduced the data so well, but there it is.

The referee writes:

  Further, the manuscript assumes that the apparent size-period
  distribution of false positives can be described by a single
  two-dimensional normal distributions and suggest that this is
  consistent with background eclipsing binary stars being the dominant
  source of false positives among Kepler planet candidates. However,
  previous works have demonstrated that both hierarchical triple stars
  including an eclipsing binary and large planets orbiting the fainter
  member of a binary star are both significant sources of false
  positives. Further, both the rate of false positives and the relative
  abundance of the types of false positives are functions of orbital
  period and planet candidate size. This analysis completely ignores
  these important issues.

Of the four sentences in this paragraph, three contain major errors.
First, our manuscript describes the size-period distribution of false
positives by a linearly-varying function of size and period (c.f. Eq.
(6)), not as a "single two-dimensional normal distribution
[sic]"---perhaps this is a typo? Second, in the third sentence the
referee seems to be complaining that we do not allow for size and
period variations in the background distribution, but it is explicit
in both the text and Figure 1(!) that the distribution of background
contamination we use varies over the size-period plane. Third, in the
final sentence the referee claims that we ignore these issues (related
to contamination), but we do not; for example, we note in the
Conclusion with respect to Petigura et al (2013) and Foreman-Mackey et
al (2014) that "[in contrast to our work] Neither study attempted to
account for contamination from falsely-identified candidate transit
events, controlling this instead through careful choice of threshold."
Similar statements appear throughout the paper. It is our view that
incorporating contamination in the analysis is an important advance
from our paper. In the remaining sentence of this paragraph, the
referee notes that many types of contamination are present in the
data, whose relative importance varies across the size-period plane;
we agree, but note that our contamination model is purely empirical
and agnostic to the source of contamination. While we claim that our
fit is consistent with the amount of contamination from background
eclipsing binaries determined from careful studies, nothing in our
model is based on the detailed properties of this particular source of
contamination.  We have added several sentences throughout the paper
that emphasise this point.  Finally, we have become aware though
helpful comments from Tim Morton that Morton & Johnson (2011) presents
an early measurement of contamination in Kepler, and we now cite that
paper in addition to Fressin et al. (2013) where appropriate.

The referee writes:

  The analysis is based entirely on a publicly available planet
  candidate list, which is heterogeneous, but these issues are not
  addressed. Further, they do not provide enough information about the
  data set used to allow someone else to repeat their calculation.

We state in the introduction that "The Q17 data release describes 2598
“candidate” planetary transit signals identified by the Kepler team
from observations of stars in the “EX” observing program (which are
primarily G-type main-sequence stars similar to our own Sun (Batalha
et al. 2010)), giving the inferred planetary period and radius for
each." and, following Eq. (1) discussing our fiducial SNR, that "In
our analysis, we obtain these quantities from the Kepler Input Catalog
(Batalha et al. 2010; Brown et al. 2011) and the MAST Kepler
archive[web link]." Additionally, we have released all our code for
the analysis at https://github.com/ farr/kepler-selection (linked in
the paper's conclusion). This includes the code that processes the
data from the above-mentioned sources, code that performs the fitting,
and code that produces the plots found in the paper. We are not sure
what additional information the referee would require to repeat our
calculation. We also note that, to our knowledge, no other estimate of
eta-Earth has been performed using this data set (the result from
Petigura et al. 2013 and the following Foreman-Mackey et al. 2014 used
the same lightcurves from Kepler, but a customised pipeline). In some
senses, this data set is heterogeneous (F, G, K stars), but in many
important ways it is not: same instrument (Kepler), same pipeline,
etc. We stand by our claim that this measurement refers to planets
around Sun-like stars, though, unlike Petigura et al. 2013 and
Foreman-Mackey et al. 2014, we did not restrict to G-type stars.

The referee writes:

  Additionally, they do not model the observational uncertainties in
  the estimated planet radii and summarily dismiss a superior
  statistical analysis which both accounts for this effect and
  demonstrates the importance of this effect. The claim that their
  result for eta_Earth is more precise than Foreman-Mackey et al. 2014
  is at best misleading. It reflects a serious misunderstanding of the
  main points from that paper. I am confident that this manuscript's
  estimated uncertainty is smaller, because they ignore important
  effects, extrapolate using a less flexible model, and make additional
  unjustified assumptions.

We do not attempt to model uncertainties in the estimates of period
and radius; Foreman-Mackey et al 2014 do not attempt to account for
contamination. Neither of these choices reflects a mis-understanding
(in fact, Foreman-Mackey et al 2014 devote a paragraph of their
discussion section to "False Positives," noting that their analysis
does not account for them and proposing an extension of their work to
examine their effect), but rather reflects the difficulty of
constructing models that simultaneously reflect both effects. We have
added several sentences to the conclusion section to make this point
clearer. It is true that including period-radius uncertainty might
increase the range of our constraints on eta-Earth and other
parameters, but not by very much; note that we have tested our
sensitivity to uncertainty, stating in Section 2 "In spite of several
candidates with very large uncertainties in measured parameters, we
have found that our fit is essentially unchanged when applied to
synthetic observations with periods and radii re-drawn from the range
of observational uncertainties quoted in the Q17 data release."  We
re-iterate this point now in the conclusion.

The referee writes:

  I could go on with many more effects that are not adequately
  modeled, and often not even discussed. But it's not a referee's job
  to develop a research plan for them.

  In summary, the manuscript presents one basic calculation for an
  interesting, but complex problem which deserves a more detailed
  analysis. I can understand journals accepting one or two early
  analyses that are overly simplistic. But by now the field has plenty
  of simplistic estimates. To make a significant contribution, the field
  needs people to devote the time required to perform a more complete
  analysis that pays attention to both astronomical and statistical
  issues. Such a paper is almost certainly longer than appropriate for
  ApJL. But there is certainly room for at least several more analyses
  that build on each other to develop a progressively more complete
  model to understand the frequency and distribution of small planets
  around other stars.

We disagree with the referee that our calculation is "basic" or
"overly simplistic." If it were basic to incorporate contamination or
empirically estimate selection effects then I am sure that previous
papers would have incorporated elements of our analysis. Instead, we
feel that our novel statistical methods for approaching this problem,
applied to a previously-unexamined data set provide a useful
independent estimate of the size-period distribution of exoplanets
that confirms some aspects of previous analyses and calls others into
question. It is true that our model of the distribution and selection
functions is simple---but it fits the data extremely well. Should we
complicate it in spite of this?  The referee seems unusually hostile
to our modelling choices; given the large number of errors in the
referee's statements about our paper, we would like to request a
second referee's opinion on our paper.

### 2015-03-10 paper/kepler-selection.tex
Cleanup of references and re-format properly for ApJL.
emulateapj apparently doesn't include the proper bibliography style files, so
added hapj.
Changed Ilya's e-mail.
Replaced non-printed \and with \& in author list; testing git and Latex installation on new laptop. :)
Proposed changes for response to referee report.

@@
 We find $\etaearth = \earange$ (90\% CL).  We conclude that each star
-hosts $\rplrange$ planets with $P \lesssim 3 \mathrm{yr}$ and $R
-\gtrsim 0.2 \REarth$, that the peak of the planet radius distribution
-lies at $\Rpeak = \rpeakrange \REarth$, and that $\ln P$ and $\ln R$
-are correlated with correlation coefficient $r = \corrcoeffrange$ (all
-90\% CL).  Our empirical model for false-positive contamination is
-consistent with the dominant source being background eclipsing binary
-stars \citep{Fressin2013}, with $\fposrange$ (90\% CL) of the
+hosts $\rplrange$ planets with $P \lesssim 3 \mathrm{yr}$ and
+$R \gtrsim 0.2 \REarth$, that the peak of the planet radius
+distribution lies at $\Rpeak = \rpeakrange \REarth$, and that $\ln P$
+and $\ln R$ are correlated with correlation coefficient
+$r = \corrcoeffrange$ (all 90\% CL).  Our empirical model for
+false-positive contamination does not depend on the detailed
+properties of the source of contamination, but is consistent with the
+dominant source being background eclipsing binary stars
+\citep{Morton2011,Fressin2013}, with $\fposrange$ (90\% CL) of the
 candidates being false-positives.  The distribution of planets we
@@
       and density of background contamination.} (Top) The inferred
-    detection probability versus signal-to-noise ratio (see
-    Eq.\ \eqref{eq:pdetect}) from our parameterised model of selection
+    detection probability versus signal-to-noise ratio (see Eq.\
+    \eqref{eq:pdetect}) from our parameterised model of selection
     effects.  The solid line is the posterior median detection
@@
     candidate period and large radius, consistent with the dominant
-    source of contamination being background eclipsing
-    binaries \citep{Fressin2013}.  Overall, our model finds
-    $\fposrange$ of the candidates are false-positive background
-    signals, consistent with the analysis in \citet{Fressin2013}.}
+    source of contamination being background eclipsing binaries
+    \citep{Morton2011,Fressin2013} (but note that our model does not
+    depend on the details of the sources contamination).  Overall, our
+    model finds $\fposrange$ of the candidates are false-positive
+    background signals, consistent with the analysis in
+    \citet{Morton2011,Fressin2013}.}
 \end{figure}
@@
 \end{multline}
-where $\Delta \ln P = \ln P_\mathrm{max} - \ln P_\mathrm{min}$, $\ln
-P_\mathrm{mid} = 1/2\left(\ln P_\mathrm{max} - \ln P_\mathrm{min}
-\right)$, $\Delta \ln R = \ln R_\mathrm{max} - \ln R_\mathrm{min}$,
+where $\Delta \ln P = \ln P_\mathrm{max} - \ln P_\mathrm{min}$,
+$\ln P_\mathrm{mid} = 1/2\left(\ln P_\mathrm{max} - \ln P_\mathrm{min}
+\right)$,
+$\Delta \ln R = \ln R_\mathrm{max} - \ln R_\mathrm{min}$,
 $\ln R_\mathrm{mid} = 1/2\left(\ln R_\mathrm{max} - \ln R_\mathrm{min}
-\right)$.  $\Nbg$, the expected number of background false-positive
-events; $P_\mathrm{max}$, $P_\mathrm{min}$, $R_\mathrm{max}$, and
+\right)$.
+$\Nbg$, the expected number of background false-positive events;
+$P_\mathrm{max}$, $P_\mathrm{min}$, $R_\mathrm{max}$, and
 $R_\mathrm{min}$, the boundaries in the $P$-$R$ plane within which
@@
 density of background events, are parameters of our model.  This is a
-purely empirical model for the background contamination, but is
-reasonable if the chief contaminant is background eclipsing binaries
-\citep{Fressin2013,Duquennoy1991}.  The posterior on the background
-number density in the $P$-$R$ plane appears in
-Figure\ \ref{fig:det-bg}.
+purely empirical model for the background contamination, not dependent
+on any properties of the source of contamination, but is reasonable if
+the chief contaminant is background eclipsing binaries
+\citep{Morton2011,Fressin2013,Duquennoy1991}.  The posterior on the
+background number density in the $P$-$R$ plane appears in Figure\
+\ref{fig:det-bg}.
 
@@
 $\fposrange$ (90\% CL), consistent with previous work
-\citep{Fressin2013} estimating the contamination in the Kepler
-candidate set.  Our model has the peak of the planet period-radius
-distribution at $\Rpeak = \rpeakrange \REarth$, $P_\mathrm{peak} =
-\ppeakrange \mathrm{yr}$, and the distribution of planetary radii and
-periods is correlated, with correlation coefficient $r =
-\corrcoeffrange$ (all at 90\% CL).
+\citep{Morton2011,Fressin2013} estimating the contamination in the
+Kepler candidate set.  Our model has the peak of the planet
+period-radius distribution at $\Rpeak = \rpeakrange \REarth$,
+$P_\mathrm{peak} = \ppeakrange \mathrm{yr}$, and the distribution of
+planetary radii and periods is correlated, with correlation
+coefficient $r = \corrcoeffrange$ (all at 90\% CL).
 
@@
     contamination rate is consistent with previous work
-    \citep{Fressin2013}. (Lower Right) The planet number density per
-    logarithmic planet period.  The density peaks at $P = \ppeakrange
-    \mathrm{yr}$ (90\% CL). }
+    \citep{Morton2011,Fressin2013}. (Lower Right) The planet number
+    density per logarithmic planet period.  The density peaks at $P =
+    \ppeakrange \mathrm{yr}$ (90\% CL). }
 \end{figure}
@@
 \citep{Petigura2013} from a subset of the available Kepler data and a
-customised pipeline to search for transit signals.  They both
-accounted for selection effects by measuring the recoverability of
-synthetic transit signals injected into their data, in contrast to our
-approach of empirically determining them from the observed data.
-Neither study attempted to account for contamination from
-falsely-identified candidate transit events, controlling this instead
-through careful choice of threshold.  Both studies used a more
-flexible model for the intrinsic distribution of planets than ours.
-Our result for $\etaearth$ is consistent with, but more precise than,
-\citet{Foreman-Mackey2014} and (somewhat) inconsistent with
-\citet{Petigura2013}.
+customised pipeline to search for transit signals.  The official
+Kepler pipeline used to produce the candidate list used in this work
+is an independent code that searches the same lightcurves for planets
+using different algorithms. Both \citet{Petigura2013} and
+\citet{Foreman-Mackey2014} accounted for selection effects by
+measuring the recoverability of synthetic transit signals injected
+into their data, in contrast to our approach of empirically
+determining them from the observed data.  Neither study attempted to
+account for contamination from falsely-identified candidate transit
+events as we do, controlling this instead through careful choice of
+threshold.  Both studies used a more flexible model for the intrinsic
+distribution of planets than ours, and \citet{Foreman-Mackey2014}
+accounted for the observational uncertainty in periods and radii
+explicitly in their model.  \citet{Foreman-Mackey2014} found that the
+discrepancy between their results and those of \citet{Petigura2013}
+was likely due to the un-modelled uncertainty in period and radius in
+the latter work.  Our result for $\etaearth$ is consistent with, but
+more precise than, \citet{Foreman-Mackey2014} and inconsistent with
+\citet{Petigura2013}.  We have checked that our results remain largely
+unchanged when re-analysing synthetic data sets that contain periods
+and radii drawn from the range of observational uncertainties quoted
+by the Kepler pipeline.  Combining the methods of
+\citet{Foreman-Mackey2014} for treating measurement uncertainty with
+the methods of this work for dealing with contamination is left for
+future work.
 
@@
 
+We thank Timothy D. Morton for helpful comments on this manuscript.
 The code implementing this analysis is available under an open-source
@@
 
+\bibliographystyle{hapj}
 \bibliography{kepler-selection}

### 2015-03-11 paper/response.txt
Ilya's changes to the response temper my enraged excesses.

@@
 As you will see from our response to the referee attached, we have
-identified a number of major errors in the referee's discussion of our
-work. Additionally, the referee seems unusually hostile to our
-modelling choices (e.g. "I could go on with many more effects that are
-not adequately modeled, and often not even discussed. But it's not a
+identified a number of errors in the referee's discussion of our work.
+Additionally, the referee seems unusually hostile to our modelling
+choices (e.g. "I could go on with many more effects that are not
+adequately modeled, and often not even discussed. But it's not a
 referee's job to develop a research plan for them."). As you will see
@@
 Given these issues, could we request to have a second, independent
-referee comment on our revised work?  We don't know the procedures for
-this, but if appropriate we would be happy for both referees to see
-this note to the editor and the attached referee response.
+referee comment on our revised work? If appropriate we would be happy
+for both referees to see this note to the editor and the attached
+referee response.
 
@@
 
-There are several problems with the referee's main points of criticism
-of our paper, and also several outright mistakes. Nevertheless, in
-some cases we feel that the referee has valid criticisms of our paper.
-Below we explain the changes we have made to our manuscript to attempt
-to address the referee's criticism and where we feel that criticism is
-in error.
+The referee has made several valid criticisms of our paper, which we
+address below. Nevertheless, we feel that there are some
+misunderstandings in the referee's main points of criticism. Below we
+explain the changes we have made to our manuscript to attempt to
+address the referee's criticism and where we feel that criticism is in
+error.
 
@@
 
-Of the four sentences in this paragraph, three contain major errors.
 First, our manuscript describes the size-period distribution of false
 positives by a linearly-varying function of size and period (c.f. Eq.
-(6)), not as a "single two-dimensional normal distribution
-[sic]"---perhaps this is a typo? Second, in the third sentence the
-referee seems to be complaining that we do not allow for size and
-period variations in the background distribution, but it is explicit
-in both the text and Figure 1(!) that the distribution of background
-contamination we use varies over the size-period plane. Third, in the
-final sentence the referee claims that we ignore these issues (related
-to contamination), but we do not; for example, we note in the
-Conclusion with respect to Petigura et al (2013) and Foreman-Mackey et
-al (2014) that "[in contrast to our work] Neither study attempted to
-account for contamination from falsely-identified candidate transit
-events, controlling this instead through careful choice of threshold."
-Similar statements appear throughout the paper. It is our view that
+(6)). Second, in the third sentence the referee seems to be
+complaining that we do not allow for size and period variations in the
+background distribution, but it is explicit in both the text and
+Figure 1 that the distribution of background contamination we use
+varies over the size-period plane. Third, in the final sentence the
+referee claims that we ignore these issues (related to contamination),
+but we do not; for example, we note in the Conclusion with respect to
+Petigura et al (2013) and Foreman-Mackey et al (2014) that "[in
+contrast to our work] Neither study attempted to account for
+contamination from falsely-identified candidate transit events,
+controlling this instead through careful choice of threshold." Similar
+statements appear throughout the paper. It is our view that
 incorporating contamination in the analysis is an important advance
-from our paper. In the remaining sentence of this paragraph, the
-referee notes that many types of contamination are present in the
-data, whose relative importance varies across the size-period plane;
-we agree, but note that our contamination model is purely empirical
-and agnostic to the source of contamination. While we claim that our
-fit is consistent with the amount of contamination from background
-eclipsing binaries determined from careful studies, nothing in our
-model is based on the detailed properties of this particular source of
-contamination.  We have added several sentences throughout the paper
-that emphasise this point.  Finally, we have become aware though
-helpful comments from Tim Morton that Morton & Johnson (2011) presents
-an early measurement of contamination in Kepler, and we now cite that
-paper in addition to Fressin et al. (2013) where appropriate.
+from our paper. The referee notes that many types of contamination are
+present in the data, whose relative importance varies across the
+size-period plane; we agree, but note that our contamination model is
+purely empirical and agnostic to the source of contamination. While we
+claim that our fit is consistent with the amount of contamination from
+background eclipsing binaries determined from careful studies, nothing
+in our model is based on the detailed properties of this particular
+source of contamination. We have added several sentences throughout
+the paper that emphasise this point. Finally, we have become aware
+though helpful comments from Tim Morton that Morton & Johnson (2011)
+presents an early measurement of contamination in Kepler, and we now
+cite that paper in addition to Fressin et al. (2013) where
+appropriate.
 
@@
 We state in the introduction that "The Q17 data release describes 2598
-“candidate” planetary transit signals identified by the Kepler team
-from observations of stars in the “EX” observing program (which are
+"candidate" planetary transit signals identified by the Kepler team
+from observations of stars in the "EX" observing program (which are
 primarily G-type main-sequence stars similar to our own Sun (Batalha
@@
 archive[web link]." Additionally, we have released all our code for
-the analysis at https://github.com/ farr/kepler-selection (linked in
+the analysis at https://github.com/farr/kepler-selection (linked in
 the paper's conclusion). This includes the code that processes the
@@
 Petigura et al. 2013 and the following Foreman-Mackey et al. 2014 used
-the same lightcurves from Kepler, but a customised pipeline). In some
-senses, this data set is heterogeneous (F, G, K stars), but in many
-important ways it is not: same instrument (Kepler), same pipeline,
-etc. We stand by our claim that this measurement refers to planets
-around Sun-like stars, though, unlike Petigura et al. 2013 and
-Foreman-Mackey et al. 2014, we did not restrict to G-type stars.
+the some of the same lightcurves from Kepler, but a customised
+pipeline). In some senses, this data set is heterogeneous (F, G, K
+stars), but in many important ways it is not: same instrument
+(Kepler), same pipeline, etc.
 
@@
 We disagree with the referee that our calculation is "basic" or
-"overly simplistic." If it were basic to incorporate contamination or
-empirically estimate selection effects then I am sure that previous
-papers would have incorporated elements of our analysis. Instead, we
-feel that our novel statistical methods for approaching this problem,
-applied to a previously-unexamined data set provide a useful
-independent estimate of the size-period distribution of exoplanets
-that confirms some aspects of previous analyses and calls others into
-question. It is true that our model of the distribution and selection
-functions is simple---but it fits the data extremely well. Should we
-complicate it in spite of this?  The referee seems unusually hostile
-to our modelling choices; given the large number of errors in the
-referee's statements about our paper, we would like to request a
-second referee's opinion on our paper.
+"overly simplistic." Our approach to incorporating contamination and
+emprically estimating selection effects is based on novel statistical
+methods and represents a step forward relative to previous analyses.
+The application of these methods to a previously-unexamined data set
+provides a useful independent estimate of the size-period distribution
+of exoplanets that confirms some aspects of previous analyses and
+calls others into question. The fact that our simple model of the
+distribution and selection functions fits the data extremely well
+indicates that, at present, the data quality and quantity are not
+sufficient to warrant more complicated modeling.
